---
title: "Statistical analysis of VL data"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
---

```{r load-packages, echo = FALSE, message=F,output=F}
knitr::opts_chunk$set(echo = F)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(GGally)
library(cowplot)
library(emmeans)
library(drtmle)
library(SuperLearner)
library(nloptr)
library(janitor)
theme_set(theme_bw() + theme(legend.position = "top", panel.grid.minor = element_blank()))

source(here("R", "directory-funs.R"))

#stopifnot(dir.exists("/Volumes/trials/"))

vl_lims = c(1, 7)
ic_cats = c("< 1 mcg/mL", "1-3 mcg/mL", "> 3 mcg/mL")
vl_title = expression(paste("Viral load (log"[10], " copies/mL)"))
ep_labels = tibble(
  levels = c("fp", "peak", "nadir", "mean"),
  label = c("First positive", "Peak", "Nadir", "Mean")
)

rx_labeller = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

```


```{r load-data}

outcome_map = tribble(
  ~trt_ic80, ~trt_ic80_code,
  "Placebo, IC80 < 1 mcg/mL", 1,
  "Placebo, IC80 >= 1 mcg/mL", 2,
  "VRC01 (10mg/kg), IC80 < 1 mcg/mL", 3,
  "VRC01 (10mg/kg), IC80 >= 1 mcg/mL", 4,
  "VRC01 (30mg/kg), IC80 < 1 mcg/mL", 5,
  "VRC01 (30mg/kg), IC80 >= 1 mcg/mL", 6
)

outcome_pool_map = tribble(
  ~trt_ic80_pool, ~trt_ic80_pool_code,
  "Placebo, IC80 < 1 mcg/mL", 1,
  "Placebo, IC80 >= 1 mcg/mL", 2,
  "VRC01 (Pooled), IC80 < 1 mcg/mL", 3,
  "VRC01 (Pooled), IC80 >= 1 mcg/mL", 4
)

vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols()) %>%
  mutate(
    ic80gte1 = factor(gmt80ls >= 1, levels = c(F, T), labels = c("IC80 < 1 mcg/mL", "IC80 >= 1 mcg/mL")),
    rx_lab = factor(rx_code2, levels = rx_labeller$rx_code2, labels = rx_labeller$trt_cat),
    rx_pool = if_else(rx_code2 == "C", "Placebo", "VRC01 (Pooled)"),
    trt_ic80 = paste(rx_lab, ic80gte1, sep = ", "),
    trt_ic80_pool = paste(rx_pool, ic80gte1, sep = ", ")
  ) %>%
  left_join(outcome_map, by = "trt_ic80") %>%
  left_join(outcome_pool_map, by = "trt_ic80_pool")

first_positive = vl_data %>% 
  dplyr::filter(days_fp == 0) %>%
  mutate(
    isSA703 = study_region_code == 1,
    isNSA703 = study_region_code == 2,
    isSwiss704 = study_region_code == 3
  ) %>%
  arrange(trt_ic80_code)


first_positive %>%
  tabyl(trt_ic80, trt_ic80_code) %>%
  kable() %>%
  kable_styling(full_width= F)

first_positive %>%
  tabyl(trt_ic80_pool, trt_ic80_pool_code) %>%
  kable() %>%
  kable_styling(full_width= F)

```

## Overview

See methods-discussion.md

## First positive analysis


```{r}

ggplot(first_positive, aes(x = protocol, y = log10vl)) +
  geom_boxplot() +
  geom_point() +
  facet_grid(ic80gte1 ~ rx_lab) +
  labs(y = "First positive log10 viral load", x = "Protocol")

```


```{r}

ggplot(first_positive, aes(x = study_region_cat, y = log10vl)) +
  geom_boxplot() +
  geom_point() +
  facet_grid(ic80gte1 ~ rx_lab) +
  labs(y = "First positive log10 viral load") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### All groups

```{r fp-drtmle}
# South America/704  is biggest category

outcome = first_positive$log10vl
trt_groups = first_positive$trt_ic80_code
cov = select(first_positive, isSA703, isNSA703, isSwiss704) %>% 
  mutate(across(.fns = as.numeric))

drtmle_fit_fp = drtmle(Y = outcome, A = trt_groups, W = cov,
             stratify = FALSE,
             SL_g = c("SL.glm", "SL.glm.interaction", "SL.gam"),
             SL_Q = c("SL.glm", "SL.glm.interaction", "SL.gam"),
             #glm_Qr = "gn",
             #glm_gr = "Qn",
             SL_gr = c("SL.glm", "SL.mean"),
             SL_Qr = c("SL.glm", "SL.mean"),
             cvFolds = 10,
             SL_method = "method.NNLS", family = gaussian(), returnModels = TRUE)

        
# compare point estimates for all different estimators
drtmle_fit_fp_ci = ci(drtmle_fit_fp, est = c("drtmle", "tmle", "aiptw", "gcomp"))
drtmle_mm_fp = map2_df(drtmle_fit_fp_ci, names(drtmle_fit_fp_ci),
                       ~mutate(as_tibble(.x, rownames = "trt_ic80_code"), model = .y)) %>%
  mutate(trt_ic80_code = as.numeric(trt_ic80_code)) %>%
  left_join(outcome_map, by = "trt_ic80_code")

```


```{r fp-lm}

mod_fp_interaction = lm(log10vl ~ study_region_cat + trt_ic80, data = first_positive)

summary(mod_fp_interaction)

anova(mod_fp_interaction) %>% as_tibble(rownames = "Variable") %>%
  kable(caption = "ANOVA: Dose group with IC80 >= 1 interaction", digits = 3) %>%
  kable_styling(full_width = T)

emmeans(mod_fp_interaction, ~ trt_ic80) %>%
  as_tibble() %>%
  rename(est = emmean,  cil =  lower.CL,  ciu = upper.CL) %>%
  mutate(model = "LM") %>%
  bind_rows(drtmle_mm_fp) %>%
  ggplot(aes(x = trt_ic80, y = est, ymin = cil, ymax = ciu, colour = model)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  labs(y = "First positive log10 viral load", x = "VRC01 Groups") +
  coord_flip()

emmeans(mod_fp_interaction, pairwise ~ ic80gte1*rx_lab, adjust = "mvt") 

```

### Pooled Treatment

```{r fp-drtmle-pool}
# South America/704  is biggest category

outcome = first_positive$log10vl
trt_groups_pool = first_positive$trt_ic80_pool_code
cov = select(first_positive, isSA703, isNSA703, isSwiss704) %>% 
  mutate(across(.fns = as.numeric))

drtmle_fit_fp_pool = drtmle(Y = outcome, A = trt_groups_pool, W = cov,
             stratify = FALSE,
             SL_g = c("SL.glm", "SL.glm.interaction", "SL.gam"),
             SL_Q = c("SL.glm", "SL.glm.interaction", "SL.gam"),
             #glm_Qr = "gn",
             #glm_gr = "Qn",
             SL_gr = c("SL.glm", "SL.mean"),
             SL_Qr = c("SL.glm", "SL.mean"),
             cvFolds = 10,
             SL_method = "method.NNLS", family = gaussian(), returnModels = TRUE)

# compare point estimates for all different estimators
drtmle_fp_pool_ci = ci(drtmle_fit_fp_pool, est = c("drtmle", "tmle", "aiptw", "gcomp"))
drtmle_mm_fp_pool = map2_df(drtmle_fp_pool_ci, names(drtmle_fp_pool_ci),
                       ~mutate(as_tibble(.x, rownames = "trt_ic80_pool_code"), model = .y)) %>%
  mutate(trt_ic80_pool_code = as.numeric(trt_ic80_pool_code)) %>%
  left_join(outcome_pool_map, by = "trt_ic80_pool_code")

```


```{r fp-lm}
mod_fp_pool_interaction = lm(log10vl ~ study_region_cat + trt_ic80_pool, data = first_positive)

summary(mod_fp_pool_interaction)

anova(mod_fp_pool_interaction) %>% as_tibble(rownames = "Variable") %>%
  kable(caption = "ANOVA: Dose group with IC80 >= 1 interaction", digits = 3) %>%
  kable_styling(full_width = T)

emmeans(mod_fp_pool_interaction, ~ trt_ic80_pool) %>%
  as_tibble() %>%
  rename(est = emmean,  cil =  lower.CL,  ciu = upper.CL) %>%
  mutate(model = "LM") %>%
  bind_rows(drtmle_mm_fp_pool) %>%
  ggplot(aes(x = trt_ic80_pool, y = est, ymin = cil, ymax = ciu, colour = model)) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  labs(y = "First positive log10 viral load", x = "VRC01 Groups") +
  coord_flip()

emmeans(mod_fp_interaction, pairwise ~ ic80gte1*rx_lab, adjust = "mvt") 

```

```{r fp-pool}

mod_fp_pool_main_effects = lm(log10vl ~ study_region_cat + rx_pool * ic80gte1, data = first_positive)

summary(mod_fp_pool_interaction)

anova(mod_fp_pool_interaction) %>% as_tibble(rownames = "Variable") %>%
  kable(caption = "ANOVA: Dose group with IC80 >= 1 interaction", digits = 3) %>%
  kable_styling(full_width = T)

emmeans(mod_fp_pool_interaction, ~ ic80gte1+rx_pool) %>%
  as_tibble() %>%
  ggplot(aes(x = rx_pool, y = emmean, ymin = lower.CL, ymax = upper.CL)) +
  geom_pointrange() +
  facet_wrap(~ic80gte1, ncol = 1) +
  labs(y = "First positive log10 viral load (95% CI)", x = "VRC01 Groups") +
  coord_flip()

emmeans(mod_fp_pool_interaction, pairwise ~ ic80gte1*rx_pool, adjust = "none")$contrasts %>%
  kable(digits = 2) %>%
  kable_styling()


emmeans(mod_fp_pool_interaction, pairwise ~ ic80gte1*rx_pool, adjust = "none")$contrasts %>% confint %>% as_tibble() %>%
  ggplot(aes(x = contrast, y = estimate, ymin = lower.CL, ymax = upper.CL)) +
  geom_pointrange() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(y = "Difference in positive log10 viral load (95% CI)", x = "") +
  coord_flip()

```

**everything below is old, due to copying and pasting of document placeholder**

```{r final-adata}

vl_data_analysis_full = vl_data %>%
  dplyr::filter(!pub_id %in% low_sample_ptids & on_art == 0 & has_neut_data) %>%
  left_join(dplyr::select(neut_data, -contains("cat")), by = "pub_id") %>%
  mutate(log10VL = log10(vl_num),
         rx_lab = factor(rx_code2, levels = rx_lab$rx_code2, labels = rx_lab$trt_cat))

 write_csv(dplyr::select(vl_data_analysis_full, -flag, -has_neut_data, -infection_pre80wks, -rx),
                         "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/amp-vl-neut-all.csv")

vl_data_analysis = vl_data_analysis_full %>%
  dplyr::filter(days_fp >= 0) %>%
  mutate(ic80ls_cat = factor(if_else(gmt80ls < 1, 0, 1),
                           levels = 0:1,
                           labels = c(expression(paste("IC80 < 1 mcg/mL")), 
                                      expression(paste("IC80" >= "1 mcg/mL")))))
  
neut_data_long = neut_data %>%
  dplyr::select(-contains("cat")) %>%
  gather("key", "value", -pub_id, -nisolates) %>%
  mutate(key = str_replace_all(key, "0", "0_")) %>%
  separate(key, into = c("titer", "isolate"), sep = "_") %>%
  pivot_wider(names_from = "titer", values_from = "value")


kinetics_summary = vl_data_analysis %>%
  group_by(pub_id, rx_lab, protocol, nisolates, across(contains("gmt"))) %>%
  nest() %>%
  summarize(
    peak = map(data, ~tibble(days_fp = .$days_fp[which.max(.$log10VL)],
                             log10VL = .$log10VL[which.max(.$log10VL)])),
    nadir = map(data, ~tibble(days_fp = .$days_fp[which.min(.$log10VL)],
                             log10VL = .$log10VL[which.min(.$log10VL)])),
    fp = map(data, ~tibble(days_fp = 0,
                             log10VL = .$log10VL[.$days_fp == 0])),
    mean = map(data, ~tibble(days_fp = NA_real_,
                             log10VL = mean(.$log10VL)))
    ,`.groups` = "drop") %>%
  pivot_longer(names_to = "endpoint", values_to = "tib", c(peak, nadir, fp, mean)) %>%
  unnest(c(tib)) %>%
  mutate(endpoint_plot = factor(endpoint, levels = ep_labels$levels, labels = ep_labels$label))
  

censored_pub = subset(vl_data_analysis, cens == 1) %>%
  distinct(pub_id) %>% pull(pub_id)

```


## Exploratory data analysis 

### Longitudinal

```{r vl-scatter-peak, fig.height=7,fig.width=6.5}

ggplot(vl_data_analysis, aes(x = days_fp, y = log10VL, group = pub_id, colour = rx_lab)) +
  geom_line(alpha = 0.5) +
  geom_point(data = dplyr::filter(kinetics_summary, endpoint == "peak"), alpha  = 0.75) +
  scale_y_continuous(breaks = 2:7) +
  labs(color = "", x = "Days post first positive", y = vl_title) +
  facet_wrap(~protocol, nrow = 2) 


```


```{r vl-scatter-groups}

ggplot(vl_data_analysis, aes(x = days_fp, y = log10VL, group = pub_id, colour = rx_lab)) +
  geom_line(alpha = 0.5) +
  #geom_point(data = dplyr::filter(kinetics_summary, endpoint == "peak"), alpha  = 0.75) +
  scale_y_continuous(breaks = 2:7) +
  labs(color = "", x = "Days post first positive", y = vl_title) +
  facet_wrap(~ic80ls_cat+rx_lab, nrow = 2, labeller = labeller(ic80ls_cat = label_parsed)) +
  theme(legend.position = "none")


```


```{r vl-time-peak}

kinetics_summary %>% 
  dplyr::filter(endpoint == "peak") %>%
  ggplot(aes(x = days_fp, y = log10VL, colour = rx_lab)) +
  geom_point(alpha  = 0.5) +
  scale_y_continuous(limits = vl_lims, breaks = 1:7) +
  labs(color = "", x = "Days post first positive", y = "Peak log10VL") +
  facet_wrap(~protocol, nrow = 2)

```

```{r vl-time-nadir}

kinetics_summary %>%
  dplyr::filter(endpoint == "nadir") %>%
  ggplot(aes(x = days_fp, y = log10VL, colour = rx_lab)) +
  geom_point(alpha  = 0.5) +
  scale_y_continuous(limits = vl_lims, breaks = 1:7) +
  labs(color = "", x = "Days post first positive", y = "Nadir log10VL") +
  facet_wrap(~protocol, nrow = 2)

```

### Kinetic Summary

```{r kinetics-bp}

 bquote("IC80 >= 1 mcg/mL")

kinetics_summary %>%
  mutate(rx_lab2 = str_replace(rx_lab, " ", "\n"),
         ic80_cat = factor(if_else(gmt80ls < 1, 0, 1),
                           levels = 0:1,
                           labels = c(expression(paste("IC80 < 1 mcg/mL")), 
                                      expression(paste("IC80" >= "1 mcg/mL"))))
         )  %>%
  ggplot(aes(x = rx_lab2, y = log10VL, colour = rx_lab2)) +
  geom_boxplot(width = 0.3) +
  scale_y_continuous(breaks = 2:7) +
  geom_point()+
  facet_grid(cols = vars(ic80_cat), rows = vars(endpoint_plot), switch = "y", labeller = labeller(ic80_cat = label_parsed)) +
  labs(y = vl_title, x = "") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        legend.position = "none",
        axis.text.x = element_text(size = 9)) 

```

### Neutralization summary

The IC50 and IC80 data is very well correlated within isolate class (most frequent, most sensitive, least sensitive). And generally in agreement in resistance category. Two outliers standout because the IC80 suggests resistance while the IC50 suggests none.

Using an orthogonal regression, the IC50 and IC80 data track well together. The intercept of 3.3-fold shift between IC50 and IC80 suggest that the IC50 and IC80 are closer together than a Hill-slope of 1 (which would correspond to 4-fold change). The slope of 1 (if IC50 goes up x-fold, then IC80 goes up x-fold) implies that the hill slope is conserved over the range of sensitive viruses in these data. 

```{r}

denning_reg = function(formula, dat){
  pca <- prcomp(formula, dat)
  tibble(slp = with(pca, rotation[2,1] / rotation[1,1]),
         int = with(pca, center[2] - slp*center[1])
  )
}

pls = map(c("ms", "mf", "ls"), function(i){
  tmp_dat = neut_data_long %>%
    dplyr::filter(isolate == i & gmt80 < 100) %>%
    mutate(log10gmt50 = log10(gmt50), log10gmt80 = log10(gmt80))
  
  den_lm = denning_reg(~log10gmt50 + log10gmt80, tmp_dat)
  
  ggplot(data = tmp_dat, aes(x = log10gmt50, y = log10gmt80)) + 
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = log10(3), ymax = Inf, fill = "red", alpha = 0.5) +
    annotate("rect", xmin = log10(3/3.33), xmax = Inf, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.5) +
    annotate("rect", xmin = -Inf, xmax = log10(1/3.33), ymin = -Inf, ymax = log10(1), fill = "blue", alpha = 0.5) +
    #geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +  
    geom_point() +
    geom_abline(slope=den_lm$slp, intercept=den_lm$int, color="blue") +
    annotate("text", x = -0.5, y = 1.75, label = paste0('slp=', round(den_lm$slp,2),
                                                      "; ratio=", round(10^den_lm$int,2))) +
    geom_hline(yintercept = log10(c(1, 3)), linetype = "dashed") +
    ggtitle(i) #+
    #annotate("text", x = 1, y = 200, label = linear(lm(fo, tmp_dat)), 
    #                      colour="black", size = 3, parse=TRUE)
  
})

plot_grid(plotlist = pls)
 
```

```{r}


vl_data_analysis %>%
  dplyr::select(contains("gmt"), nisolates) %>%
  dplyr::filter(nisolates > 1) %>%
  ggpairs(columns = 1:6, mapping = ggplot2::aes(colour=factor(nisolates)))
  
  
```


## Supplementary figures

```{r vl-scatter-cens}

vl_data_analysis %>%
  dplyr::filter(pub_id %in% censored_pub) %>%
  ggplot(aes(x = days_fp, y = log10VL, group = pub_id, colour = rx_lab)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha  = 0.5) +
  scale_y_continuous(limits = vl_lims, breaks = 1:7) +
  facet_wrap(~protocol, nrow = 2)

```

```{r kinetics-protocol-bp}

kinetics_summary %>%
  mutate(rx_lab2 = str_replace(rx_lab, " ", "\n")) %>%
  ggplot(aes(x = rx_lab2, y = log10VL, colour = rx_lab2)) +
  geom_boxplot(width = 0.3) +
  scale_y_continuous(breaks = 2:7) +
  geom_point()+
  facet_grid(cols = vars(protocol), rows = vars(endpoint_plot), switch = "y") +
  labs(y = vl_title, x = "") +
  theme(strip.placement = "outside", strip.background = element_blank(),
        legend.position = "none",
        axis.text.x = element_text(size = 9)) 

```

