---
title: "Create mlx viral load data"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---


# Creating MLX-VL data

Monolix has some required specifications and variables it is expecting so this file generally creates datasets that meet those requirements. 

This script focuses on creating data for fitting just VL models, which requires simpler data than the PKPD models. For the PKPD models, the data must be processed with a "VL dose", the infection times, and other fixed parameters as necessary. 

- Key inputs for the data:
  - Data generated from `analysis/data-processing/02_data-processing`
    - processed vl data with neut data merged on: `data/processed-data/adata-vl.csv`
      - neut/demo data from here
      
This will create two datasets: all participants (158 participants) and placebo only (62 participants).

The processing is minimal, essentially setting up the limit variable for censoring (-1 is upper bound censored, 1 is below lloq):
`limit = if_else(cens == 1, "0", if_else(cens == -1, "1e9", "."))`

```{r load-packages, message=F,output=F,warning=F}

knitr::opts_chunk$set(echo = T)
library(tidyverse)
library(here)
library(testthat)

theme_set(theme_bw() + theme(legend.position = "top"))

options(scipen = 999)

```


```{r load-funs-data}
source(here("R", "directory-funs.R"))
```
    
    
```{r load-data}
adata_vl = read_csv(clean_data_here("adata-vl.csv"), col_types = cols())

```

```{r mlx-setup, class.source = "fold-show"}

amp_vl_all = adata_vl %>%
  mutate(
    limit = if_else(cens == 1, "0", if_else(cens == -1, "1e9", "."))
    ) %>%
  dplyr::select(pub_id, TIME = days_enrollment, DV = vl, 
                DV_log = log10vl, fp_day, days_fp, tpf_est = fp_post_tpf, cens, limit,
                 protocol, study_region_cat, contains("gmt"), nisolates, study, rx_code2)

placebo_vl = filter(amp_vl_all, rx_code2 == "C") 

```

```{r error-checks}

expect_equal(n_distinct(amp_vl_all$pub_id), 158)
expect_equal(all(!is.na(amp_vl_all)), TRUE)
expect_equal(n_distinct(placebo_vl$pub_id), 62)

```

```{r mlx-vl-save, class.source = "fold-show"}

write_csv(amp_vl_all, mlx_data_here("amp-mlx-vl.csv"))

write_csv(placebo_vl, mlx_data_here("amp-mlx-vl-placebo.csv"))

```