---
title: "Survival Analysis for Various Time-based Outcomes"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

Several curves are compared here.

The main incidence curves of interest are:
  - time between last infusion pre first positive to first positive
  - time from first positive to ART initiation
  
The main comparisons are:
  - 703 vs 704
  - resistant vs sensitive
  - across treatment groups
  - sensitive+trt groups

```{r load-packages, echo = FALSE, message=F,output=F,warning=F}
knitr::opts_chunk$set(echo = T)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(GGally)
library(cowplot)
library(survival)
library(survminer)
library(knitr)
theme_set(theme_bw() + theme(legend.position = "top", panel.grid.minor = element_blank()))

source(here("R", "directory-funs.R"))

rx_lab = tibble(
  levels = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

rx_lab_pool = tibble(
  levels = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

time_lab_inf = "Weeks post final pre-RNA+ infusion"
time_lab_fp = "Weeks post first RNA+"
#colors = c('gray','light green','forest green')
colors = c('gray40','steelblue','navy')
color_protocol = c("magenta", "purple")

colors_grp = c("gray", "#2CA02C")
ic80gte1_colors = c("#1F77B4", "coral")
full_colors = c('navy',"#D62728", "#1F77B4", 'coral')

ic80gte1_colors_named = c(`IC80 < 1 mcg/mL` = "#1F77B4", `IC80 >= 1 mcg/mL` = "coral")

ic80_label = c(`IC80 < 1 mcg/mL` =  bquote("#1F77B4"),
               `IC80 >= 1 mcg/mL` =  "coral"
             )

ytitle_fp = "Cumulative fraction first RNA+"
ytitle_art = "Cumulative fraction ART initiation"
opts_chunk$set(dev = c("png", "pdf"))

clean_surv_test = function(surv_test){
  as_tibble(surv_test$p.value, rownames = "grp1") %>% 
    gather("grp2", "value", -grp1) %>%
    unite(col = "comparison", grp1, grp2, sep = " vs. ") %>%
    rename(lrt_pvalue = value)

}

```

```{r load-data}

time_summary_data = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols()) %>%
  mutate(
    protocol_full = if_else(protocol == "HVTN 703", "HVTN 703/HPTN 081", "HVTN 704/HPTN 085"),
    grp = rx_lab,
    rx_pool = if_else(rx_code2 == "C", "Placebo", "VRC01 (Pooled)"),
    ic80gte1 = factor(gmt80ls >= 1, levels = c(F, T), 
                      labels = c("IC80 < 1 mcg/mL", "IC80 >= 1 mcg/mL")),
     trt_ic80_pool = paste(rx_pool, ic80gte1, sep = ", ")
    )

```

```{r art-data}

art_surv_data = time_summary_data %>%
  dplyr::filter() %>%
  mutate(
    any_art = !is.na(artstartdt),
    time_to_art = if_else(any_art, art_day - fp_day, final_day - fp_day)/7
    )

stopifnot(all(art_surv_data$time_to_art > 0))
stopifnot(all(!is.na(art_surv_data$any_art)))

```

```{r fp-data}

stopifnot(nrow(subset(time_summary_data, is.na(final_infusion_day))) == 0)
stopifnot(all(!is.na(time_summary_data$fp_day) & time_summary_data$fp_day < Inf))
stopifnot(all(!is.na(time_summary_data$final_infusion_pre_fp_day) & time_summary_data$final_infusion_pre_fp_day < Inf))
stopifnot(!any(is.na(time_summary_data$fp_day)))

fp_surv_data = time_summary_data %>%
  dplyr::filter() %>%
  mutate(
    time_to_fp = (fp_day - final_infusion_pre_fp_day)/7
    )

```

## ART Incidence Curves

```{r sample-sizes}

art_surv_data %>%
  group_by(protocol_full, grp) %>%
  summarize(N = n_distinct(pub_id), .groups = "drop") %>%
  pivot_wider(values_from = N, names_from = grp) %>%
  janitor::adorn_totals() %>%
  kable() %>%
  kable_styling(full_width = F)
```

```{r sample-size-endpoints}

art_surv_data %>%
  group_by(protocol_full, any_art, grp) %>%
  summarize(N = n_distinct(pub_id), `.groups` = "drop") %>%
  pivot_wider(values_from = N, names_from = grp) %>%
  janitor::adorn_totals() %>%
  kable() %>%
  kable_styling(full_width = F)

```


```{r trt, fig.height=7}

surv_trt = survfit(Surv(time_to_art, any_art) ~ grp, data = art_surv_data)
names(surv_trt$strata) <- gsub("grp=", "", names(surv_trt$strata)) 
# pairwise_survdiff(Surv(time_to_art, any_art) ~ rx_code2, data = art_surv_data) 
# loc_lr_pvalues = as_tibble(round(ada_surv_loc_test[["p.value"]], 2), rownames = " ") %>%
#   replace_na(list(`Left Arm` = " "))

surv_trt_test = pairwise_survdiff(Surv(time_to_art, any_art) ~ grp, 
                 data = art_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

ggsurvplot(
  surv_trt,
  data = art_surv_data,
  legend.title = "",
  risk.table = T,
  palette = colors,
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event"
) 

surv_trt_test

```

```{r amp703-art-tbl, fig.height=7}

surv_703 = survfit(Surv(time_to_art, any_art) ~ grp, 
                   data = subset(art_surv_data, protocol == "HVTN 703"))
names(surv_703$strata) <- gsub("grp=", "", names(surv_703$strata)) 


surv_703_test = pairwise_survdiff(Surv(time_to_art, any_art) ~ grp, 
                                  data = subset(art_surv_data, protocol == "HVTN 703"), 
                                  p.adjust.method = "none") %>%
  clean_surv_test() %>%
  mutate(comparison = paste0("703: ", comparison))

pl_703 = ggsurvplot(
  surv_703,  
  risk.table = T,
  palette = colors,
  data = subset(art_surv_data, protocol == "HVTN 703"),
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
) +
  ggtitle("HVTN 703/HPTN 081")

pl_703

```

```{r amp704-art-tbl, fig.height=7}

surv_704 = survfit(Surv(time_to_art, any_art) ~ grp, 
                   data = subset(art_surv_data, protocol == "HVTN 704"))
names(surv_704$strata) <- gsub("grp=", "", names(surv_704$strata)) 


surv_704_test = pairwise_survdiff(Surv(time_to_art, any_art) ~ grp, 
                                  data = subset(art_surv_data, protocol == "HVTN 704"), 
                                  p.adjust.method = "none") %>%
  clean_surv_test() %>%
  mutate(comparison = paste0("704: ", comparison))

pl_704 = ggsurvplot(
  surv_704,
  risk.table = T,
  data = subset(art_surv_data, protocol == "HVTN 704"),
  palette = colors,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
) +
  ggtitle("HVTN 704/HPTN 085")

pl_704
```

```{r inc-art-trt}

plot_grid(nrow = 2, rel_heights = c(12,2),
          plot_grid(pl_703$plot + theme(legend.position = "none"), 
                    pl_704$plot + theme(legend.position = "none")),
          get_legend(pl_703$plot)
)

```

```{r art-protocol-tbl}

surv_protocol = survfit(Surv(time_to_art, any_art) ~ protocol_full, 
                   data = art_surv_data)
names(surv_protocol$strata) <- gsub("protocol_full=", "", names(surv_protocol$strata)) 

surv_procotol_test = pairwise_survdiff(Surv(time_to_art, any_art) ~ protocol_full, 
                                       data = art_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

pl_protocol = ggsurvplot(
  surv_protocol,
  risk.table = T,
  data = art_surv_data,
  palette = color_protocol,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
) 

pl_protocol

```

```{r art-trt-pool}


surv_trt_pool = survfit(Surv(time_to_art, any_art) ~ rx_pool, 
                   data =art_surv_data)
names(surv_trt_pool$strata) <- gsub("rx_pool=", "", names(surv_trt_pool$strata)) 

surv_trt_pool_test =  pairwise_survdiff(Surv(time_to_art, any_art) ~ rx_pool, 
                  data = art_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

pl_trt_pool = ggsurvplot(
  surv_trt_pool,  
  risk.table = T,
  palette = colors_grp,
  data = art_surv_data,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
)

pl_trt_pool

```

```{r art-ic80-tbl}

surv_ic80 = survfit(Surv(time_to_art, any_art) ~ ic80gte1, 
                   data = art_surv_data)
names(surv_ic80$strata) <- str_remove(names(surv_ic80$strata), "ic80gte1=") 

surv_ic80_test = pairwise_survdiff(Surv(time_to_art, any_art) ~ ic80gte1, 
                                   data = art_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

pl_ic80 = ggsurvplot(
  surv_ic80,
  risk.table = T,
  data = art_surv_data,
  palette = ic80gte1_colors,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
) 

pl_ic80

```

```{r art-trt-ic80}

surv_full= survfit(Surv(time_to_art, any_art) ~ trt_ic80_pool, 
                   data = art_surv_data)
names(surv_full$strata) <- str_remove(names(surv_full$strata), "trt_ic80_pool=") 

surv_full_test =pairwise_survdiff(Surv(time_to_art, any_art) ~ trt_ic80_pool, 
                                   data = art_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

pl_full = ggsurvplot(
  surv_full,
  risk.table = T,
  data = art_surv_data,
  palette = full_colors,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_fp,
  fun = "event",
  ylim = c(0, 1)
) 

pl_full

```


```{r art-tests}

art_tests = surv_trt_test %>%
  bind_rows(surv_703_test) %>%
  bind_rows(surv_704_test) %>%
  bind_rows(surv_ic80_test) %>%
  bind_rows(surv_trt_pool_test) %>%
  bind_rows(surv_full_test) %>%
  mutate(test = "ART initiation (log-rank)") %>%
  select(test, everything()) %>%
  filter(!is.na(lrt_pvalue))

art_tests %>%
  kable(digits = 3) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(1, valign = "top")

```

## Last infusion (pre positive RNA) to first positive

```{r fp-pooled, fig.height=7}

fp_surv_pool = survfit(Surv(time_to_fp) ~ grp, data = fp_surv_data)
names(fp_surv_pool$strata) <- gsub("grp=", "", names(fp_surv_pool$strata)) 
fp_surv_trt_test = pairwise_survdiff(Surv(time_to_fp) ~ grp, 
                                     data = fp_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

ggsurvplot(
  fp_surv_pool,
  data = fp_surv_data,
  legend.title = "",
  risk.table = T,
  palette = colors,
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event"
) 

```

```{r fp-amp703, fig.height=7}

fp_surv_703 = survfit(Surv(time_to_fp) ~ grp, 
                   data = subset(fp_surv_data, protocol == "HVTN 703"))
names(fp_surv_703$strata) <- gsub("grp=", "", names(fp_surv_703$strata)) 

fp_surv_703_test = pairwise_survdiff(Surv(time_to_fp) ~ grp, 
                                     data = subset(fp_surv_data, protocol == "HVTN 703"), 
                                     p.adjust.method = "none") %>%
  clean_surv_test() %>%
  mutate(comparison = paste0("703: ", comparison))

fp_pl_703 = ggsurvplot(
  fp_surv_703,  
  risk.table = T,
  palette = colors,
  data = subset(fp_surv_data, protocol == "HVTN 703"),
  legend.title = "",
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event",
  ylim = c(0, 1),
  font.x = c(12),
  xlim = c(0, 80),
  break.x.by = 20
) +
  ggtitle("HVTN 703/HPTN 081")

fp_pl_703

```

```{r fp-amp704, fig.height=7}

fp_surv_704 = survfit(Surv(time_to_fp) ~ grp, 
                   data = subset(fp_surv_data, protocol == "HVTN 704"))
names(fp_surv_704$strata) <- gsub("grp=", "", names(fp_surv_704$strata)) 

fp_surv_704_test = pairwise_survdiff(Surv(time_to_fp) ~ grp, 
                                     data = subset(fp_surv_data, protocol == "HVTN 704"),
                                     p.adjust.method = "none") %>%
  clean_surv_test() %>%
  mutate(comparison = paste0("704: ", comparison))

fp_pl_704 = ggsurvplot(
  fp_surv_704,
  risk.table = T,
  data = subset(fp_surv_data, protocol == "HVTN 704"),
  palette = colors,
  legend.title = "",
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event",
  font.x = c(12),
  ylim = c(0, 1)
) +
  ggtitle("HVTN 704/HPTN 085")

fp_pl_704

```

```{r inc-fp-trt}

plot_grid(nrow = 2, rel_heights = c(12,2),
          plot_grid(fp_pl_703$plot + theme(legend.position = "none"), 
                    fp_pl_704$plot + theme(legend.position = "none")),
          get_legend(fp_pl_704$plot)
)

```

```{r fp-protocol}

fp_surv_protocol = survfit(Surv(time_to_fp) ~ protocol_full, 
                   data = fp_surv_data)
names(fp_surv_protocol$strata) <- gsub("protocol_full=", "", names(fp_surv_protocol$strata)) 

pairwise_survdiff(Surv(time_to_fp) ~ protocol_full, data = fp_surv_data, p.adjust.method = "none")

fp_pl_protocol = ggsurvplot(
  fp_surv_protocol,
  risk.table = T,
  data = fp_surv_data,
  palette = color_protocol,
  legend.title = "",
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event",
  ylim = c(0, 1)
) 

fp_pl_protocol

```

```{r fp-trt-pool}

fp_surv_trt_pool = survfit(Surv(time_to_fp) ~ rx_pool, data = fp_surv_data)
names(fp_surv_trt_pool$strata) <- gsub("rx_pool=", "", names(fp_surv_trt_pool$strata)) 

fp_surv_trt_pool_test = pairwise_survdiff(Surv(time_to_fp) ~ rx_pool, 
                                          data = fp_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

fp_pl_trt_pool = ggsurvplot(
  fp_surv_trt_pool,  
  risk.table = T,
  palette = colors_grp,
  data = fp_surv_data,
  legend.title = "",
  ylab = ytitle_art,
  xlab = time_lab_inf,
  fun = "event",
  ylim = c(0, 1)
)

fp_pl_trt_pool

```

```{r fp-ic80}

fp_surv_ic80 = survfit(Surv(time_to_fp) ~ ic80gte1, data = fp_surv_data)
names(fp_surv_ic80$strata) <- str_remove(names(fp_surv_ic80$strata), "ic80gte1=") 

fp_surv_ic80_test = pairwise_survdiff(Surv(time_to_fp) ~ ic80gte1, 
                                          data = fp_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

fp_pl_ic80 = ggsurvplot(
  fp_surv_ic80,
  risk.table = T,
  data = fp_surv_data,
  palette = ic80gte1_colors,
  legend.title = "",
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event",
  ylim = c(0, 1)
) 

fp_pl_ic80

```


```{r fp-trt-ic80}

fp_surv_full = survfit(Surv(time_to_fp) ~ trt_ic80_pool, data = fp_surv_data)
names(fp_surv_full$strata) <- str_remove(names(fp_surv_full$strata), "trt_ic80_pool=") 

fp_surv_full_test = pairwise_survdiff(Surv(time_to_fp) ~ trt_ic80_pool, 
                                      data = fp_surv_data, p.adjust.method = "none") %>%
  clean_surv_test()

fp_pl_full = ggsurvplot(
  fp_surv_full,
  risk.table = T,
  data = fp_surv_data,
  palette = full_colors,
  legend.title = "",
  ylab = ytitle_fp,
  xlab = time_lab_inf,
  fun = "event",
  ylim = c(0, 1)
) 

fp_pl_full

```


```{r fp-tests}

fp_tests = fp_surv_trt_test %>%
  bind_rows(fp_surv_703_test) %>%
  bind_rows(fp_surv_704_test) %>%
  bind_rows(fp_surv_ic80_test) %>%
  bind_rows(fp_surv_trt_pool_test) %>%
  bind_rows(fp_surv_full_test) %>%
  mutate(test = "infusion-FP (log-rank)") %>%
  select(test, everything()) %>%
  filter(!is.na(lrt_pvalue))

fp_tests %>%
  kable(digits = 3) %>%
  kable_styling(full_width = F) %>%
  collapse_rows(1, valign = "top")

```

# Manuscript Supp Figure

```{r lrt-tests}

art_tests %>%
  bind_rows(fp_tests) %>%
  write_csv("incidence-analysis_files/figure-html/lrt_tests.csv")

```


```{r supp-time-to-event, fig.height=7, fig.width=10, dev = c("pdf", "png")}

make_panel = function(survA, survB, 
                      leg_size = 9, 
                      text_size = 9,
                      title_size = 9){
  legend = get_legend(survA$plot + 
                        theme(
                          legend.background = element_blank(),
                          legend.text = element_text(size = leg_size))
                      )
  
  pl_theme = theme(legend.position = "none", 
                   axis.text.y = element_text(size = text_size),
                   axis.text.x = element_text(size = text_size),
                   axis.title.y = element_text(size = title_size),
                   axis.title.x = element_text(size = title_size)
                   )
  
  plot_grid(
    plot_grid(survA$plot + pl_theme, survB$plot + pl_theme, nrow = 1),
    legend, 
    nrow = 2, rel_heights = c(12, 2)
  )
}

plot_grid(make_panel(fp_pl_protocol, pl_protocol), 
          make_panel(fp_pl_trt_pool, pl_trt_pool), 
          make_panel(fp_pl_ic80, pl_ic80), 
          make_panel(fp_pl_full+guides(color=guide_legend(nrow=2,byrow=TRUE)), 
                                       pl_full, leg_size = 8),
          nrow = 2, ncol = 2,
          labels = LETTERS[1:4])

```
