---
title: "PK, Neutralization, and first positive VLs - Model result analysis"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

## Summary

- There is little evidence for cases where infection happened before an infusion happened. This is different than an infusion after a first positive, which happens frequently. The former could latently alter first positives without affecting infection risk (i.e., titer at infection < titer at first positive).

- In counter-factual simulations where VRC01 was turned off. We see an increase in VLs measured at the matched observation time for sensitive viruses.
   - The increase did not reach first positive levels for the resistant group: the PKPD suppression does not sufficiently explain the full difference in VL measurements.
   - There were no cases of potentially detecting an earlier first positive with VRC01 off. There were only two people with observation times (one each) between infection and first positive where predicted VL was >0 but below detection.
  
## Setup

This has some odds and ends analysis for PK, neut, and VL. Dan has developed the dose-response models for first positive.

The main question here is the relationship of first positive to time of infection and last infusion. This depends on having the infection time estimated from the VL model analysis.

Basically, at time of infection, the titer is make or break on protection. The threshold identified for protection here is a PT80 in the 300 range. However, for VL modulation, we find the titer must be higher. The paradox is, if higher titers prevention infection vs. modulation, how do we have post-acquisition titers high enough to modulate VL. The AMP study does have post-exposure infusions though that could explain.

From the data, one approach is assess titers at infection vs. titers near first positive. If post-exposure infusion explains modulation, we would expect maybe higher titers near first positive. This is challenging though because there is lag effect: high titers suppress VL prior to first positive and the first positive may be lower because of higher titers prior to the measurement vs. simultaneously with measurement. Titers are subject to PK, with non-linear effects on VL that are unobserved compounded by sparse sampling.

 - Tally cases of infusions between infection time and first positive
 - Duration duration from previous infusion to first positive, vs. first positive.

A model-based approach: Re-simulate PKPD-V in VRC01 participants stopping VRC01 at infection time. Then predict VL at the first positive measurement time to compare with the observed first positive.

```{r, message=F, warning = F}
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(cowplot)
library(janitor)
library(cowplot)
library(GGally)
library(knitr)
source(here("R", "directory-funs.R"))
source(here("R", "rxode-funs.R"))
source(here("R", "mlx-res-processing.R"))
source(here("models", "rxode-models", "rxode-pkpd-models.R"))
theme_set(theme_bw())


vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols()) 

vrc01_dose = read_csv(mlx_data_here("amp-mlx-pkpd-indirect-holteT0.csv"), col_types = cols()) %>%
  filter(ADM == 1) %>%
  distinct(pub_id, TIME, AMT, RATE) %>%
  mutate(AMT = as.numeric(AMT), RATE = as.numeric(RATE)) %>%
  arrange(pub_id, TIME) %>%
  group_by(pub_id) %>%
  mutate(DOSENO = row_number()) %>%
  ungroup()

vl_mod_summary_all = read_csv(clean_data_here("final_vl_summary.csv"), col_types = cols()) 

vl_mod_vrc01 = vl_mod_summary_all %>% filter(rx_code2 != "C")
vl_mod_placebo = vl_mod_summary_all %>% filter(rx_code2 == "C")

vl_mod_vrc01_pred_raw =  get_indiv_preds("PKPD/adj-pkpd-model-no-omega") %>%
  rename(pub_id = ID)

# confirmed that ipred = sim pos
vl_sims = read_csv(clean_data_here("amp-individual-vrc01-sims.csv"), col_types = cols()) %>%
  left_join(vl_mod_vrc01_pred_raw, by = c("pub_id", "days_enrollment" = "time"))

vl_mod_vrc01_preds = vl_mod_vrc01_pred_raw %>%
  left_join(select(vl_mod_vrc01, pub_id, fp_day), by = "pub_id") %>%
  filter(time == fp_day) %>%
  mutate(first_pos_ipred = log10(indivPredMode)) %>%
  select(pub_id, first_pos_ipred)

pk_sim_pre_fp = read_csv(here("output/pk-modeling/sim-pk-withVLneut.csv"), col_types = cols()) %>%
  filter(days_enrollment <= fp_day) %>%
  left_join(select(vl_mod_vrc01, pub_id, model_infection_day), by = "pub_id") %>%
  mutate(pt80 = centr_conc/gmt80ls, sens = gmt80ls <= 1)

# 2 vrc01 ptids only have first positive, werent in model
stopifnot(nrow(subset(pk_sim_pre_fp, round(model_infection_day) == days_enrollment)) == 96)

```

## Counter-factual FP modeling

Turning off VRC01 altogether explains some of the reduction in VL but not all of it.

```{r}
vl_mod_vrc01_pred_raw %>%
  ggplot(aes(x = log10(indivPredMode), y = log10(DV))) +
  geom_point() +
  geom_abline()


```

Without VRC01, were there any increases in VL that would've been caught earlier?
No (2 observation times, all below log10 LLoQ of 1.3).

```{r}
vl_sims %>%
  filter(days_enrollment > model_infection_day & days_enrollment <= fp_day & !first_positive_row & !is.na(DV)) %>% 
  select(pub_id, sim_type, log10V)
```

```{r}

vl_sims %>%
  dplyr::filter(first_positive_row & sim_type %in% c("NoTreatment", "VRC01")) %>%
  pivot_wider(id_cols = pub_id, names_from = sim_type, values_from = log10V) %>%
  select(-pub_id) %>%
  ggpairs()

```

```{r}

vl_sims %>%
  filter(first_positive_row & sim_type %in% c("NoTreatment", "VRC01")) %>%
  ggplot(aes(x = sim_type, y = log10V)) +
  geom_boxplot() +
  geom_point() +
  geom_path(aes(group = pub_id), alpha = 0.5) +
  facet_wrap(~ic80_cat) +
  labs(x = "", y = "first pos VL")

```

## Looking at PK data

Looking at the data among sensitive viruses, only a few questionable cases where VRC01 infusion happened after infection but before first positive. 

```{r}
pk_sim_pre_fp %>%
  filter(gmt80ls <= 1 & (days_enrollment - fp_day) < 60) %>%
  ggplot(aes(x = (days_enrollment - fp_day)/7, y = pmax(1, pt80))) +
  geom_line() +
  scale_y_log10() +
  geom_vline(aes(xintercept = (model_infection_day - fp_day)/7)) +
  geom_vline(aes(xintercept = (tpf_day - fp_day)/7), colour = "red") +
  facet_wrap(~pub_id) 


```

Titers are all within an order of magnitude even in cases where titer might be higher at first positive time.

```{r}

titer_summary = pk_sim_pre_fp %>%
  filter(!is.na(model_infection_day)) %>%
  group_by(pub_id) %>%
  mutate(
    fp_vl = log10vl[first_positive],
    fp_pt80 = pt80[days_enrollment == fp_day],
    infection_pt80_pf = pt80[days_enrollment == tpf_day],
    infection_pt80_mod = pt80[days_enrollment == model_infection_day] 
    ) %>%
  distinct(pub_id, fp_vl, fp_pt80, infection_pt80_pf, infection_pt80_mod, gmt80ls, sens) 


titer_summary %>%
  ggplot(aes(x = pmax(1, fp_pt80), y = pmax(1,infection_pt80_mod))) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline() +
  facet_wrap(~sens) +
  labs(y = "pt80 at nlme infection time")

titer_summary %>%
  ggplot(aes(x = pmax(1, fp_pt80), y = pmax(1,infection_pt80_pf))) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline() +
  facet_wrap(~sens) +
  labs(y = "pt80 at paul infection time")

```

Timing is close on either end. This is interesting from a prevention perspective, but has little impact on the ODE predictions in a deterministic setting.

```{r}

ggplot(vl_mod_vrc01, aes(x = (fp_day - final_infusion_pre_fp_day)/7, y = first_pos)) +
  geom_point() +
  labs(y = "first pos VL", x = "Weeks btw first positive and previous infusion")


ggplot(vl_mod_vrc01, aes(x = (tpf_day - final_infusion_pre_fp_day)/7, y = first_pos)) +
  geom_point() +
  labs(y = "first pos VL", x = "Weeks btw pred. acquisition (PF) and final infusion (pre-fp)")

ggplot(vl_mod_vrc01, aes(x = (model_infection_day - final_infusion_pre_fp_day)/7, y = first_pos)) +
  geom_point() +
  labs(y = "first pos VL", x = "Weeks btw pred. acquisition (PKPD-V) and final infusion (pre-fp, negative VL)")

```
