---
title: "PK-Modeling"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
---

```{r load-packages, echo = FALSE, message=F,output=F}
knitr::opts_chunk$set(echo = F)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(lixoftConnectors)
library(RxODE)

theme_set(theme_bw() + theme(legend.position = "top"))

options(scipen = 999)

initializeLixoftConnectors(software = "monolix")

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

save_shared_folder = F

```


```{r load-funs-data}

source(here("R", "directory-funs.R"))
source(here("R", "mlx-model-funs.R"))
source(here("R", "fix-parm-funs.R"))
source(here("R", "mlx-res-processing.R"))

rx_dat = read_csv(raw_data_here("rx_dat.csv"), col_types = cols())

pk_data_nm = read_csv(mlx_data_here("pk-nm-data.csv"), col_types = cols())
median_weight = pk_data_nm %>% distinct(pub_id, weight) %>% summarize(median(weight)) %>% pull()
pk_data_analysis = read_csv(clean_data_here("adata-pk.csv"), col_types = cols())
analysis_ptids = unique(pk_data_analysis$pub_id)

vl_data_analysis = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols())

neut_data = read_csv(clean_data_here("amp-neut-blinded.csv"), col_types = cols())

pk_dir_tag = function(x) paste0("PK/", x)
pk_model_here = function(x, ...) mlx_model_here(pk_dir_tag(x), ...)

```

# Overview

Here, we just use Lily's case-control PK model, "refit" the model to get the individual parameters to allow us to draw trajectories and link to important time points like infection time.

# Refit PK

```{r model-setup}

# check_data("pk-nm-data.csv")

pk_data_types = tribble(
    ~var_name, ~type,
    "ID", "ignore",
    "AMT", "amount",
    "DV", "observation",
    "RATE", "rate",
    "TIME", "time",
    "dose", "ignore",
    "AVISITN", "ignore",
    "DOSENO", "ignore",
    "weight", "contcov",
    "study", "catcov",
    "pub_id", "id"
  )

```

```{r refit-pk, eval = F}

amp_recipe(
  model_text_file = "infusion_2cpt_ClV1QV2.txt",
  data_file = "pk-nm-data.csv",
  variable_types = pk_data_types,
  initial_estimates = FALSE,
  error_model = "proportional",
  obs_dist = "normal"
)

fix_pk_parms_cc()

runScenario()
save_project(pk_dir_tag("VRC01AMP_Case-Cntrl"))
pop_ests = get_pop_ests(pk_dir_tag("VRC01AMP_Case-Cntrl"))
pop_ests %>% # this fixes a bug related to gpmx if i want to use it
  mutate(
    value = NA_real_,
    se_sa = NA_real_,
    rse_sa = NA_real_
    ) %>%
 write_csv(pk_model_here("VRC01AMP_Case-Cntrl/populationParameters.txt"))
# write_csv(pop_ests, here(mlx_model_here(pk_loc("Rerun-Lily")), "populationParameters_raw.txt"))
```


# Explore PK

```{r}
loadProject(mlx_model_here(pk_dir_tag("VRC01AMP_Case-Cntrl"), mlxtran = T))

```

## Setup

```{r load-pk-res}

# getIndividualParameterModel()$formula
# log(Cl) = log(Cl_pop) + beta_Cl_tWT*tWT + beta_Cl_study_1*[study = 1]
# log(V2) = log(V2_pop) + beta_V2_study_1*[study = 1]

time_data = read_csv(clean_data_here("ptid-time-data.csv"), col_types = cols())
pop_ests = get_pop_ests(pk_dir_tag("VRC01AMP_Case-Cntrl"))

pop_wide = spread(pop_ests, key = parameter, value = value)

indiv_parms_no_followup = pk_data_analysis %>%
  dplyr::filter(no_followup) %>%
  transmute(
    pub_id = pub_id,
    Cl = exp(log(pop_wide$Cl_pop) + pop_wide$beta_Cl_tWT*log(weight/67.3) + pop_wide$beta_Cl_study_1*(study == 1)),
    V1 = pop_wide$V1_pop,
    Q = pop_wide$Q_pop,
    V2 = exp(log(pop_wide$V2_pop) + pop_wide$beta_V2_study_1 * (study == 1)),
    est_method = 'population - missing followup' 
  )

indiv_parms_mode = get_indiv_parms(pk_dir_tag("VRC01AMP_Case-Cntrl")) %>%
  dplyr::select(pub_id = id, contains("_mode")) %>%
  rename_with(~str_replace(., "_mode", "")) %>%
  dplyr::filter(pub_id %in% analysis_ptids) %>%
  mutate(est_method = 'mode') %>%
  bind_rows(indiv_parms_no_followup) %>%
  left_join(rx_dat, by = "pub_id") %>%
  mutate(rx_lab = factor(rx_code2, levels = rx_lab$rx_code2, labels = rx_lab$trt_cat)) %>%
  dplyr::select(-rx) %>%
  write_csv(here("output/pk-modeling", "pk-mode-parms.csv"))
 stopifnot(n_distinct(indiv_parms_mode$pub_id) == 96)

indiv_fits = get_indiv_fits(pk_dir_tag("VRC01AMP_Case-Cntrl")) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids) 

indiv_preds = get_indiv_preds(pk_dir_tag("VRC01AMP_Case-Cntrl")) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids)

# the 15 people with no concentration data,  of course are missing

missing_ptids_parms = setdiff(analysis_ptids, unique(indiv_parms_mode$pub_id))
n_distinct(missing_ptids_parms)

missing_ptids_fits = setdiff(analysis_ptids, unique(indiv_fits$pub_id))
n_distinct(missing_ptids_fits)

missing_ptids_preds = setdiff(analysis_ptids, unique(indiv_preds$pub_id))
n_distinct(missing_ptids_preds)
```

## Fits - mlx

```{r, fits}

#| fig.width=12, fig.height=12

pk_fits = ggplot(indiv_fits, aes(x = time/7, y = pmax(0.03, indivPredMode))) +
  geom_line() +
  geom_point(data = indiv_preds, aes(y = DV)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 9, ncol = 9) +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = tpf_day/7), colour = "red", linetype = 'dashed') +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks") +
  ggtitle("dashed = PF-infection time. solid = first positive")

pk_fits

ggsave(here("output/pk-modeling/pk_fits.pdf"), pk_fits, width = 12, height = 12)

```

## Fits - RxODE

```{r fit-nlmixr-setup}

dose_data = pk_data_nm %>%
  dplyr::filter(DV == "." & AMT != ".") %>%
  mutate(across(c(AMT, RATE), as.numeric)) %>%
  select(pub_id, TIME, AMT, RATE, DOSENO)

pk_2cmpt <- RxODE({
  kel = Cl/V1
  k12 = Q/V1
  k21 = Q/V2

  d/dt(centr) = -(k12+kel)*centr + k21 * peri
  d/dt(peri) = k12*centr - k21*peri
  
  Cc = centr/V1
  Pc = peri/V2
})

sim_pk <- function(final_day,
                   parms,
                   dose_info,
                   key_times = NULL,
                   output_grid_by = 1) {
  
  sim_times = c(seq(0, final_day, by = output_grid_by), key_times) %>%
    sort() %>%
    unique()


  ev = eventTable(time.units = "days") %>%
    add.sampling(sim_times) 
  
  for(i in dose_info$DOSENO){
    dose_iteration = dplyr::filter(dose_info, DOSENO == i)
    ev = ev %>%
      add.dosing(
      dose = dose_iteration$AMT,
      rate = dose_iteration$RATE,
      nbr.doses = 1,
      cmt = "centr",
      start.time = dose_iteration$TIME
      )
  }

  as_tibble(rxSolve(pk_2cmpt, params = parms, event = ev)) %>% 
    mutate(days_enrollment = as.numeric(time)) %>%
    dplyr::select(-time)

}

```


```{r test-dat-setup, eval = F}
test_pub_id = "704-0662"
test_set = subset(indiv_parms_mode, pub_id == test_pub_id) %>%
  select(where(is.numeric))
test_dose = subset(dose_data, pub_id == test_pub_id) 
vl_times = subset(time_data, pub_id == test_pub_id)

x = sim_pk(vl_times$final_vl_day, test_set, dose_info = test_dose, 
           key_times = c(subset(indiv_preds, pub_id == test_pub_id)$time,
                         subset(indiv_fits, pub_id == test_pub_id)$time,
                         vl_times$fp_day, vl_times$tpf_day, vl_times$dx_day)) %>%
  mutate(pub_id = test_pub_id)


ggplot(data = x, aes(x = days_enrollment/7, y = Cc)) +
  geom_line() +
  geom_line(data = subset(indiv_fits, pub_id == test_pub_id), aes(x = time/7, y = indivPredMode), colour = "red") +
  scale_y_log10(limits = c(0.1, NA)) +
  geom_vline(data = vl_times, aes(xintercept = fp_day/7)) 


subset(indiv_preds, pub_id == test_pub_id) %>%
  dplyr::select(pub_id, time, DV, indivPredMode) %>%
  left_join(dplyr::select(indiv_fits, pub_id, time, indivPredMode), by = c("pub_id", "time")) %>%
  left_join(dplyr::select(x, pub_id, time = days_enrollment, Cc), by = c("pub_id", "time")) 

```

```{r sim-pk}


sim_pk_dat = map_df(unique(subset(indiv_parms_mode, !is.na(Cl))$pub_id), function(i){
  
  # relevent data
  parms = dplyr::select(subset(indiv_parms_mode, pub_id == i), Cl, V1, Q, V2)
  dosing = subset(dose_data, pub_id == i) 
  vl_data = dplyr::select(subset(vl_data_analysis, pub_id == i & days_enrollment >= fp_day), 
                          pub_id, days_enrollment, cens, vl, log10vl)
  vl_data_times = vl_data$days_enrollment
  vl_times = subset(time_data, pub_id == i)
  obs_data = dplyr::select(subset(indiv_preds, pub_id == i), pub_id, time, obs_conc = DV)
  
  final_time = vl_times$final_vl_day
  extra_pk_vl_times = c(obs_data$time,
               subset(indiv_fits, pub_id == i)$time,
               vl_data_times, 
               vl_times$tpf_day
  )
               
  sim_pk(
    vl_times$final_vl_day,
    parms = parms,
    dose_info = dosing,
    key_times = extra_pk_vl_times
  ) %>%
    mutate(pub_id = i) %>%
    dplyr::select(pub_id, days_enrollment, centr_conc = Cc, peri_conc = Pc) %>%
    left_join(indiv_parms_mode, by = "pub_id") %>%
    left_join(obs_data, by = c("pub_id", "days_enrollment" = "time")) %>%
    left_join(vl_data, by = c("pub_id", "days_enrollment")) %>%
    mutate(
      tpf_day = vl_times$tpf_day,
      fp_day =  vl_times$fp_day,
      dx_day = vl_times$dx_day,
      post_fp = days_enrollment >= fp_day,
      vl_detected = !is.na(vl),
    ) 
  
})

any(is.na(sim_pk_dat$centr_conc))
any(is.na(sim_pk_dat$peri_conc))

```


```{r fits-rxode}

#| fig.width=12, fig.height=12

pk_fits_ode = ggplot(sim_pk_dat, aes(x = days_enrollment/7, y = pmax(0.03, centr_conc))) +
  geom_line() +
  geom_point(aes(y = obs_conc)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 10, ncol = 10) +
  geom_vline(aes(xintercept = tpf_day/7), colour = "red", linetype = 'dashed') +
  geom_vline(aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks") +
  ggtitle("dashed = PF-infection time. solid = first positive")

pk_fits_ode

ggsave(here("output/pk-modeling/pk_fits_ode.pdf"), pk_fits_ode, width = 12, height = 12)

```

## Save key output

Merge on neut data for full time series PKPD with viral load.

Also filter to infection data alone.

```{r save-pk-sims}

sim_pk_dat_neut = sim_pk_dat %>%
  left_join(dplyr::select(neut_data, -contains("cat")) , by = "pub_id") %>%
  write_csv(here("output/pk-modeling/pk-sims-vl-neut.csv"))

infection_day_conc = sim_pk_dat_neut %>%
  dplyr::filter(abs(days_enrollment - tpf_day) <1e-6) %>%
  dplyr::select(pub_id, days_enrollment, centr_conc, peri_conc, tpf_day, fp_day, contains("rx"), contains("gmt")) %>%
   write_csv(here("output/pk-modeling/infection-day-pkpd.csv"))

n_distinct(infection_day_conc)
infection_day_conc %>% group_by(pub_id) %>% summarize(total = n()) %>% filter(total != 1)

```


```{r}
read_csv(here("output/pk-modeling/infection-day-pkpd-x.csv")) %>%
  dplyr::select(pub_id, centr_conc) %>%
  full_join(select(infection_day_conc, pub_id, centr_conc), by = "pub_id") ->x

```

```{r save-shared-pk, eval = save_shared_folder}
write_csv(sim_pk_dat_neut, "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/sim-pk-withVLPD.csv")
write_csv(infection_day_conc, "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/predicted-pk-infection-day.csv")
```

# PKPD primer work

## Infection time titer and concentration

```{r conc-setup}

concentration_data = sim_pk_dat_neut %>%
  dplyr::filter(days_enrollment <= tpf_day & days_enrollment > 7) %>%
  gather(key = "compartment", value = "concentration", centr_conc, peri_conc) %>%
  group_by(pub_id, compartment, vars(contains("gmt"))) %>%
  summarize(
    trough = min(concentration),
    infection_time = concentration[days_enrollment == tpf_day],
    mean = mean(concentration),
    peak = max(concentration),
    `.groups` = "drop"
  ) 

trt_titer = concentration_data %>%
  gather("time", "concentration", trough, infection_time, mean, peak, factor_key = T) %>%
  left_join(raw_neut, by = "pub_id") %>%
  gather("name", "IC", gmt50ms, gmt50ls, gmt50mf, gmt80ms, gmt80ls, gmt80mf) %>%
  mutate(
    cat = if_else(str_detect(name, "50"), "ID50", "ID80"),
    cat2 = if_else(str_detect(name, "50"), "IC50", "IC80"),
    titer = pmax(0.01, pred_conc)/IC) 
```



```{r pred-vs-obs}

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(times <= infection_day) %>%
  group_by(pub_id) %>%
  summarize(
    max_pred = max(pred_conc, na.rm = T),
    max_obs = max(obs_conc, na.rm = T)
  ) %>%
  ggplot(aes(x = max_pred, y = max_obs)) +
  geom_point() +
  labs(x = "Max prediction [VRC01]", y = "Max observed [VRC01]") +
  geom_abline()

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(times <= infection_day) %>%
  group_by(pub_id) %>%
  summarize(
    min_pred = min(pred_conc, na.rm = T),
    min_obs = min(obs_conc, na.rm = T)
  ) %>%
  ggplot(aes(x = min_pred, y = min_obs)) +
  geom_point() +
  labs(x = "Min prediction [VRC01]", y = "Min observed [VRC01]") +
  geom_abline() 

```

```{r pred-concentration}

concentration_data %>%
  gather("time", "pred_conc", trough, infection_time, mean, peak, factor_key = T) %>%
  ggplot(aes(x = time, y = pmax(pred_conc, 0.01), colour = time)) +
  geom_boxplot() +
  geom_line(aes(group = interaction(pub_id)), alpha = 0.25, colour = "gray") +
  scale_y_log10() +
  geom_point() +
  labs(y = "Predicted concentrations up to infection", x = "") 

```

```{r conc-ic50}

trt_titer %>%
  dplyr::filter(time == "infection_time") %>%
  ggplot(aes(x = pmax(pred_conc, 0.01), y = pmax(IC, 0.01))) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(intercept = c(-1), linetype = 2) +
  geom_abline(intercept = c(-2), linetype = 2, colour = "red") +
  labs(y = "ICxx", x = "Pred. [VRC01] at infection") +
  annotate("text", x = 12, y = 12 / 100, label = "titer = 100", angle = 45, vjust = 1, colour = "red") +
  annotate("text", x = 1.2, y = 1.2 / 10, label = "titer = 10", angle = 45, vjust = 1) +
  annotate("text", x = 0.1, y = 0.1, label = "titer = 1", angle = 45, vjust = 1) +
  facet_wrap(~ name + cat2, nrow = 2, scales="free_x") 

```


```{r pred-titers, width = 9}

trt_titer %>%
  ggplot(aes(x = time, y = log10(titer), colour = time)) +
  geom_boxplot() +
  geom_line(aes(group = interaction(pub_id, cat, name)), alpha = 0.25, colour = "gray") +
  geom_point() +
  labs(y = "Predicted log10 titer up to infection", x = "") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x") +
  theme(axis.text.x = element_text(size = 8))

```

## Predicted placebo titers 

Concentration data distributions come from the treatment population (see above). We consider two pre-infection concentration distributions below: predicted concentrations at time of infection and mean concentrations across participants.

To determine "placebo" titers, we use a "bootstrap" procedure. For each placebo ptid, we use their individual IC50 and IC80s to determine two measures of titer. The first is the distribution of titers calculated over the distribution of predicted concentrations at time of infection in the treatment population (Infection titer). The second is the distribution of titers calculated over the distribution of mean concentrations (prior to infection) in the treatment population (Mean titer). For each participant, these distributions are summarized by their range and their median.

This procedure is repeated on the treatment population (Bootstrap treatment). Specifically, we de-couple the PK data from the given participant as an assessment of the method, and then compare that to the actual titer calculations with the paired PK data (Actual treatment).


```{r placebo-setup}

neut_data = raw_neut %>% mutate(trt = pub_id %in% concentration_data$pub_id)
#placebo_neuts = dplyr::filter(neut_data, !trt)

bootstrap_titer = 
  map_df(unique(trt_titer$time), function(i){
    map_df(unique(neut_data$pub_id), function(j){

      neut_data %>%
        dplyr::select(-contains("cat")) %>%
        dplyr::filter(pub_id == j) %>% 
        gather("name", "value", gmt50ms, gmt50ls, gmt50mf, gmt80ms, gmt80ls, gmt80mf) %>%
        mutate(
          time = i,
          cat = if_else(str_detect(name, "50"), "ID50", "ID80"),
          min_titer = min(concentration_data[[i]]) / value,
          mean_titer = mean(concentration_data[[i]]) / value,
          median_titer = median(concentration_data[[i]]) / value,
          max_titer = max(concentration_data[[i]]) / value
        )
    })
})

```


### Range of bootstrap titer calculations in placebo

These plots are complicated.

```{r}

bootstrap_titer %>%
  dplyr::filter(!trt & time == "infection_time") %>%
  ggplot(aes(x = pub_id, y = median_titer, ymin = median_titer, ymax = max_titer)) +
  geom_pointrange() +
  coord_flip() +
  scale_y_log10("'bootstrap' median infection titer through max titer") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x") +
  labs(x = "Placebo participants") +
  theme(axis.text.y = element_blank())


```

```{r}

bootstrap_titer %>%
  dplyr::filter(!trt & time == "mean") %>%
  ggplot(aes(x = pub_id, y = median_titer, ymin = median_titer, ymax = max_titer)) +
  geom_pointrange() +
  coord_flip() +
  scale_y_log10("'bootstrap' median mean titer through max titer") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x") +
  labs(x = "Placebo participants") +
  theme(axis.text.y = element_blank())


```



### Predicted placebo titers vs. treatment


```{r placebo-trt-inf-titer}

trt_titer %>%
  rename(median_titer = titer) %>%
  mutate(trt = 2) %>%
  bind_rows(bootstrap_titer) %>%
  dplyr::filter(time == "infection_time") %>%
  mutate(
    grp = factor(trt, levels = 0:2,
                 labels = c("Bootstrap\nplacebo",
                            "Bootstrap\ntreatment",
                            "Actual\ntreatment")
    )
  ) %>%
  ggplot(aes(x = grp, y = median_titer)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_line(aes(group = interaction(pub_id, name, cat)), alpha = 0.25) +
  scale_y_log10("Infection titer (bootstrap median)") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x")


```

```{r placebo-trt-inf-titer-dens, eval = F}

trt_titer %>%
  rename(median_titer = titer) %>%
  mutate(trt = 2) %>%
  bind_rows(bootstrap_titer) %>%
  dplyr::filter(time == "infection_time") %>%
  mutate(
    grp = factor(trt, levels = 0:2,
                 labels = c("Bootstrap\nplacebo",
                            "Bootstrap\ntreatment",
                            "Actual\ntreatment")
    )
  ) %>%
  ggplot(aes(x = median_titer, colour = grp)) +
  geom_density(bw = 0.5) +
  scale_x_log10("Infection titer (bootstrap median)") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x")


```

```{r placebo-trt-mean-titer}

trt_titer %>%
  rename(median_titer = titer) %>%
  mutate(trt = 2) %>%
  bind_rows(bootstrap_titer) %>%
  dplyr::filter(time == "mean") %>%
  mutate(
    grp = factor(trt, levels = 0:2,
                 labels = c("Bootstrap\nplacebo",
                            "Bootstrap\ntreatment",
                            "Actual\ntreatment")
    )
  ) %>%
  ggplot(aes(x = grp, y = median_titer)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_line(aes(group = interaction(pub_id, name, cat)), alpha = 0.25) +
  scale_y_log10("Mean titer (bootstrap median)") +
  facet_wrap(~ cat + name, nrow = 2, scales="free_x")


```

## Concentration vs. viral load

```{r vl-pk-ic50, fig.width=10, fig.height=10, warning =F}

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(infected & times > infection_day) %>%
  left_join(raw_neut, by = "pub_id") %>%
  gather("name", "value", gmt50mscat, gmt50lscat, gmt50mfcat) %>%
  mutate(value = factor(value, levels = neut_cat_order)) %>%
  ggplot(aes(x = log10(pmax(0.01, pred_conc)), y = log10(vl_num))) +
  geom_point() +
  geom_smooth(formula = 'y ~ x', method = "lm") +
  labs(x = "Predicted log10 concentration (post-infection)", y = "Log10 VL (post-infection)") +
  ggtitle("VRC01 vs VL by IC50 cats") +
  facet_grid(name ~ value)

```

```{r vl-pk-ic80,fig.width=10, fig.height=10, warning =F}

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(infected & times > infection_day) %>%
  left_join(raw_neut, by = "pub_id") %>%
  gather("name", "value", gmt80mscat, gmt80lscat, gmt80mfcat) %>%
  mutate(value = factor(value, levels = neut_cat_order)) %>%
  ggplot(aes(x = log10(pmax(0.01, pred_conc)), y = log10(vl_num))) +
  labs(x = "Predicted log10 concentration", y = "Log10 VL (post-infection)") +
  geom_point() +
  geom_smooth(formula = 'y ~ x', method = "lm") +
  ggtitle("VRC01 vs VL by IC80 cats") +
  facet_grid(name ~ value)

```

## Titer vs. viral load

```{r vl-titer-ic50, fig.width=10, fig.height=10, warning =F}

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(infected & times > infection_day) %>%
  left_join(raw_neut, by = "pub_id") %>%
  gather("name", "value", gmt50ms, gmt50ls, gmt50mf) %>%
  mutate(titer = pmax(0.01, pred_conc)/value) %>%
  ggplot(aes(x = log10(titer), y = log10(vl_num))) +
  geom_point() +
  geom_smooth(formula = 'y ~ x', method = "loess") +
  labs(x = "Predicted log10 ID50 titer", y = "Log10 VL (post-infection)") +
  ggtitle("ID50 titer vs VL") +
  facet_grid(name ~ .)

```

```{r vl-titer-ic80, fig.width=10, fig.height=10, warning =F}

read_csv(mlx_data_here("predicted_infection_concentration.csv"), col_types = cols()) %>%
  dplyr::filter(infected & times > infection_day) %>%
  left_join(raw_neut, by = "pub_id") %>%
  gather("name", "value", gmt80ms, gmt80ls, gmt80mf) %>%
  mutate(titer = pmax(0.01, pred_conc)/value) %>%
  ggplot(aes(x = log10(titer), y = log10(vl_num))) +
  geom_point() +
  geom_smooth(formula = 'y ~ x', method = "loess") +
  labs(x = "Predicted log10 ID80 titer", y = "Log10 VL (post-infection)") +
  ggtitle("ID80 titer vs VL") +
  facet_grid(name ~ .)

```

##  Titer vs. viral load parms

```{r, include = F}


titer_tcl_pmx = pmx_mlx(
  config = "standing",
  directory = mlx_model_here("viralload/placebo-models/trt-tcl-parms"),
  input = mlx_data_here("amp-dosevl-trt-CAT.csv"),
  id = "pub_id",
  dv = "DV",
  cats = neut_cats,
  conts = neut_covs,
  settings = pmx_settings(is.draft = FALSE)
)

```

```{r, fig.width=10, fig.height=10, warning =F}
pmx_plot_eta_conts(titer_tcl_pmx, scale_x_log10 = T,  
                   smooth = list(formula = "y ~ x"),
                   correl = list(size = 3, colour = "red"), labels = list(title = "Trt TCL"))
```

```{r, fig.width=10, fig.height=10, warning =F}

pmx_plot_eta_cats(titer_tcl_pmx, labels = list(title = "Trt TCL")) + 
   scale_x_discrete(limits = neut_cat_order) +
  theme(axis.text.x = element_text(angle = 90))

```



```{r visc-pk-method, eval = F}
library(VISCPK)
test_pub_id = "704-0662"

calc_conc = map_df("704-0662", function(pid){
  vl_time_grid = subset(vl_data_analysis, pub_id == pid)$days_enrollment
  #browser()
  times = c(subset(indiv_preds, pub_id == pid)$time, vl_time_grid, vl_times$tpf_day)
  ipred = c(subset(indiv_preds, pub_id == pid)$indivPredMode, rep(NA_real_, length(vl_time_grid)+1))
  obs = c(subset(indiv_preds, pub_id == pid)$DV, rep(NA_real_, length(vl_time_grid)+1))
  infusion_times = subset(test_dose, pub_id == pid)$TIME
  infusion_doses = subset(test_dose, pub_id == pid)$AMT
  tinf = with(subset(test_dose, pub_id == pid), convert_nmRATE(nmRATE = RATE, AMT = AMT))
   
  rate_parms = convert_2cmp_pkparms(CL = test_set$Cl, Vc = test_set$V1, Q = test_set$Q, Vp = test_set$V2)
    
  tibble(
    pub_id = pid,
    times = times,
    infection_day = vl_times$tpf_day,
    infected = if_else(times %in% vl_time_grid, T, F),
    obs_conc = obs,
    ipred = ipred,
    pred_conc = infusion_2comp(mtime = times, parms = rate_parms, 
                               dose = infusion_doses, Tinf = tinf, infusion_times = infusion_times)
  )
                    
  
  })


calc_conc2 = map_df("704-0662", function(pid){
  vl_time_grid = subset(vl_data_analysis, pub_id == pid)$days_enrollment
  #browser()
  times = c(subset(indiv_fits, pub_id == pid)$time, vl_time_grid)
  infusion_times = subset(test_dose, pub_id == pid)$TIME
  infusion_doses = subset(test_dose, pub_id == pid)$AMT
  tinf = with(subset(test_dose, pub_id == pid), convert_nmRATE(nmRATE = RATE, AMT = AMT))
   
  rate_parms = convert_2cmp_pkparms(CL = test_set$Cl, Vc = test_set$V1, Q = test_set$Q, Vp = test_set$V2)
    
  tibble(
    pub_id = pid,
    times = times,
    pred_conc = infusion_2comp(mtime = times, parms = rate_parms, 
                               dose = infusion_doses, Tinf = tinf, infusion_times = infusion_times)
  )
                    
  
  })

calc_conc3 = map_df("704-0662", function(pid){
  vl_time_grid = subset(vl_data_analysis, pub_id == pid)$days_enrollment
  #browser()
  times = seq(0, max(vl_time_grid), by = 1)
  infusion_times = subset(test_dose, pub_id == pid)$TIME
  infusion_doses = subset(test_dose, pub_id == pid)$AMT
  tinf = with(subset(test_dose, pub_id == pid), convert_nmRATE(nmRATE = RATE, AMT = AMT))
   
  rate_parms = convert_2cmp_pkparms(CL = test_set$Cl, Vc = test_set$V1, Q = test_set$Q, Vp = test_set$V2)
    
  tibble(
    pub_id = pid,
    times = times,
    pred_conc = infusion_2comp(mtime = times, parms = rate_parms, 
                               dose = infusion_doses, Tinf = tinf, infusion_times = infusion_times)
  )
                    
  
  })

```


