---
title: "Simulate PK Data"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

PK parameters fit based on case-control data (all infusions prior to first positive RNA).

Two simulations performed here:

1. "Case-Control": PK through first positive
2. "AMP PK": PK through end of study

Both datasets will have key times merged on for easier filter to concentrations at those times.

The dosing data comes from the full infusion information rather than the PK dataset, but results from both approaches are compared.
For the full infusion data, any infusion after day 0, is assumed to occur at time x.1 to offset with viral measurements that would be pre-infusion on the same day (days_enrollment_dose, see final step in data-processing).

```{r load-packages, echo = T, message=F,output=F}
knitr::opts_chunk$set(echo = T)
library(VISCfunctions)
library(zoo)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(RxODE)

theme_set(theme_bw() + theme(legend.position = "top"))

options(scipen = 999)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

save_shared_folder = T

```


```{r load-funs-data}

source(here("R", "directory-funs.R"))
source(here("R", "mlx-res-processing.R"))

#rx_dat = read_csv(raw_data_here("rx_dat.csv"), col_types = cols())
pk_data_nm = read_csv(mlx_data_here("pk-nm-data.csv"), col_types = cols())
pk_data_analysis = read_csv(clean_data_here("adata-pk.csv"), col_types = cols())
time_data = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols())
all_dose_data = read_csv(clean_data_here("dosing-data.csv"), col_types = cols())
pk_indiv_parms = read_csv(here("output/pk-modeling", "pk-mode-parms.csv"), col_types = cols())

analysis_ptids = unique(pk_data_analysis$pub_id)
stopifnot(length(analysis_ptids) == 98)

vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols())

pk_location = "PK/VRC01AMP_Case-Cntrl"
pk_model_here = function(x, ...) mlx_model_here(pk_dir_tag(x), ...)
```


```{r pk-rxode-setup}

pk_2cmpt <- RxODE({
  kel = Cl/V1
  k12 = Q/V1
  k21 = Q/V2

  d/dt(centr) = -(k12+kel)*centr + k21 * peri
  d/dt(peri) = k12*centr - k21*peri
  
  Cc = centr/V1
  Pc = peri/V2
})

#' RxODE PK Simulator - simulated individual PK
#'
#' @param final_day 
#' @param parms tibble or named vector, needs Cl, V1, V2, Q
#' @param dose_info tibble/data.frame with dose info by columns for TIME, AMT, RATE, DOSENO
#' @param key_times vector of specific times to check 
#' @param post_infusion_draw automatically adds draw after infusions for correct peak (default 0 days)
#' @param output_grid_by time grid spacing (default daily)
#'
#' @return
#'
#' @examples 
#' test_sim = sim_pk(100, parms = tibble(Cl = 0.58, V1 = 3.53, Q = 0.676, V2 = 4.75),
#'   dose_info = tibble(TIME = c(0, 49.1), AMT = c(580, 600), RATE = c(15000, 27000), DOSENO = 1:2))
#' ggplot(test_sim, aes(x = days_enrollment, y = centr)) + geom_line()
#' 
sim_pk <- function(final_day,
                   parms,
                   dose_info,
                   key_times = NULL,
                   post_infusion_draw = 0,
                   output_grid_by = 1) {
  
  # add samples right after infusions
  post_dose_times = dose_info$TIME + dose_info$AMT/dose_info$RATE + post_infusion_draw
  
  sim_times = c(seq(0, final_day, by = output_grid_by), key_times, dose_info$TIME, post_dose_times) %>%
    sort() %>%
    unique()

  ev = eventTable(time.units = "days") %>%
    add.sampling(sim_times) 
  
  for(i in dose_info$DOSENO){
    dose_iteration = dplyr::filter(dose_info, DOSENO == i)
    ev = ev %>%
      add.dosing(
      dose = dose_iteration$AMT,
      rate = dose_iteration$RATE,
      nbr.doses = 1,
      cmt = "centr",
      start.time = dose_iteration$TIME
      )
  }

  as_tibble(rxSolve(pk_2cmpt, params = parms, event = ev)) %>% 
    mutate(days_enrollment = as.numeric(time)) %>%
    dplyr::select(-time)

}

```

# Explore PK

```{r read-pk-res}
#loadProject(mlx_model_here(pk_dir_tag("VRC01AMP_Case-Cntrl"), mlxtran = T))
indiv_fits = get_indiv_fits(pk_location) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids) 

indiv_preds = get_indiv_preds(pk_location) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids)
```

## Fits - mlx

```{r, fits}
#| fig.width=12, fig.height=12

pk_fits = ggplot(indiv_fits, aes(x = time/7, y = pmax(0.03, indivPredMode))) +
  geom_line() +
  geom_point(data = indiv_preds, aes(y = DV)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 10, ncol = 9) +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = tpf_day/7), colour = "red", linetype = 'dashed') +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks") +
  ggtitle("dashed = PF-infection time. solid = first positive")

pk_fits

ggsave(here("output/pk-modeling/pk_fits.pdf"), pk_fits, width = 12, height = 12)

```


# Dose error checking

Two participants have infusions that were not in the case-control data.

The times are rounded in my cleaned version of the dose data. RATES are calculated differently, when done with SAS
there was rounding before unit conversion (differences < 1 minute).

```{r setup-doses}

dose_data_casectrl = pk_data_nm %>%
  dplyr::filter(DV == "." & AMT != "." & pub_id %in% analysis_ptids) %>%
  mutate(across(c(AMT, RATE), as.numeric)) %>%
  select(pub_id, TIME, AMT, RATE, DOSENO) %>%
  arrange(pub_id, TIME, AMT, RATE, DOSENO)

dose_data = all_dose_data %>%
  dplyr::filter(pub_id %in% analysis_ptids) %>%
  select(pub_id, TIME = days_enrollment_dose, AMT = dose_mg, RATE, DOSENO = infusion_no, days_fp) %>%
  arrange(pub_id, TIME, AMT, RATE, DOSENO)
  
dose_data_prefp = dplyr::filter(dose_data, days_fp < 0)

# for error we just compare the pre-first positive to case-control

# should be missing .x not .y
merge_test = full_join(dose_data_casectrl, dose_data_prefp, by = c("pub_id", "DOSENO"))
stopifnot(all(!is.na(merge_test$AMT.y)))
subset(merge_test, is.na(AMT.x))

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = AMT.y-AMT.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = TIME.y-TIME.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = RATE.y/RATE.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

# reduce_comp = diffdf::diffdf(base = dose_data_casectrl, 
#                              subset(select(dose_data_prefp,  -days_fp), !( (pub_id == "704-2624" & DOSENO == 9) |
#                                                           (pub_id == "704-2901" & DOSENO == 8))))

```

# Simulation A: Through first positive

Mostly for error checking, probably no need to save this data explicitly.

```{r sim-pk-fp}

sim_pk_prefp_bothdosing = map_df(unique(pk_indiv_parms$pub_id), function(i){
  
  parms = dplyr::select(subset(pk_indiv_parms, pub_id == i), Cl, V1, Q, V2)
  stopifnot(nrow(parms) == 1)
  
  dosing = subset(dose_data_prefp, pub_id == i) 
  dosing_cc = subset(dose_data_casectrl, pub_id == i) 
  
  fp_day = subset(time_data, pub_id == i)$fp_day
  
  vl_data_prefp = vl_data %>%
    filter(pub_id == i & days_enrollment <= fp_day) %>%
    dplyr::select(pub_id, days_enrollment, cens, vl, log10vl)
  # for error checking
  virus_rows = nrow(vl_data_prefp)
  
  obs_data = subset(indiv_preds, pub_id == i) %>%
    dplyr::select(pub_id, days_enrollment = time, obs_conc = DV) %>%
    mutate(days_enrollment = round(days_enrollment))
  obs_rows = nrow(obs_data)
  
  test_sim = sim_pk(fp_day, parms = parms, dose_info = dosing_cc) %>%
    mutate(temp = "case-control-dosing", pub_id = i)
       
  out = sim_pk(fp_day, parms = parms, dose_info = dosing) %>%
    mutate(temp = "dosing-data", pub_id = i) %>%
    bind_rows(test_sim) %>%
    dplyr::select(pub_id, days_enrollment, centr_conc = Cc, peri_conc = Pc, temp) %>%
    left_join(pk_indiv_parms, by = "pub_id") %>%
    full_join(vl_data_prefp, by = c("pub_id", "days_enrollment")) %>%
    full_join(obs_data, by = c("pub_id", "days_enrollment")) %>%
    mutate(fp_day = fp_day)
  
  # these just confirm the full joins worked correctly  
  stopifnot(all(!is.na(out$centr_conc)))
  stopifnot(all(!is.na(out$peri_conc)))
  stopifnot(nrow(subset(out, !is.na(log10vl) & temp == "dosing-data")) == virus_rows)
  stopifnot(nrow(subset(out, !is.na(obs_conc) & temp == "dosing-data")) == obs_rows)
  
  out
})

stopifnot(nrow(subset(sim_pk_prefp_bothdosing, temp == "dosing-data" & days_enrollment == fp_day)) == 98)

```

Missing infusion information in the case-control data makes large difference for 2 participants. Otherwise, differences between dosing data sources are minimal (plot shows just cases > 0.1 mcg/ml absolute difference).

```{r comp-fp-conc}

sim_pk_prefp_bothdosing %>%
  filter(days_enrollment == round(days_enrollment)) %>%
  select(pub_id, days_enrollment, temp, centr_conc) %>%
  pivot_wider(values_from = centr_conc, names_from = temp) %>%
  ggplot(aes(x = pub_id, y = `dosing-data` -`case-control-dosing`)) +
    geom_point() 

sim_pk_prefp_bothdosing %>%
  dplyr::filter(days_enrollment == fp_day) %>%
  select(pub_id, temp, centr_conc) %>%
  pivot_wider(values_from = centr_conc, names_from = temp) %>%
  filter(abs(`dosing-data` - `case-control-dosing`) > 0.1) %>%
  ggplot(aes(x = pub_id, y = `dosing-data` -`case-control-dosing`)) +
    geom_point() +
    ylab("First positive concentration: infusion data vs. case-control pk data (> 0.1)")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r pre-output}

sim_pk_prefp = select(subset(sim_pk_prefp_bothdosing, temp == "dosing-data"), -temp) %>%
  mutate(first_positive = days_enrollment == fp_day)

# really should just be one fp per person, with upper censoring
stopifnot(n_distinct(subset(sim_pk_prefp, first_positive)$pub_id) == 98)

stopifnot(identical(subset(sim_pk_prefp, cens != 1 & !is.na(log10vl)), 
                     subset(sim_pk_prefp, first_positive))
          )

indiv_preds %>%
  mutate(time = round(time)) %>%
  dplyr::select(pub_id, days_enrollment = time, indivPredMode) %>%
  left_join(sim_pk_prefp, by = c("pub_id", "days_enrollment")) %>%
  ggplot(aes(x = centr_conc, y = indivPredMode)) +
  geom_point() +
  scale_y_log10("Matched MLX predictions ") +
  scale_x_log10("Sim predictions (observation times)") 

ggplot(subset(sim_pk_prefp, !is.na(obs_conc)), aes(x = centr_conc, y = obs_conc)) +
  geom_point() +
  scale_y_log10("Obs. concentrations") +
  scale_x_log10("Sim predictions (observation times)") 

```

# Simulation B: All predicted PK with all VL through ART initiation.

One person has a negative VL on dx day but all have vl measurements at first positive (by definition)

```{r full-pk-vl-data}

sim_pk_dat = map_df(unique(pk_indiv_parms$pub_id), function(i){
  
  parms = dplyr::select(subset(pk_indiv_parms, pub_id == i), Cl, V1, Q, V2)
  ptid_time_demo = subset(time_data, pub_id == i) %>%
    select(pub_id, protocol, study_region_cat, contains("_day"), contains("gmt"), nisolates, vl_model_flag)
  stopifnot(nrow(parms) == 1)
  stopifnot(nrow(ptid_time_demo) == 1)
 
  dosing = subset(dose_data, pub_id == i) 
  dose_info_merge = select(dosing, pub_id, recent_infusion_time = TIME, infusion_no = DOSENO) %>%
    mutate(days_enrollment = recent_infusion_time)

  vl_data_ptid = vl_data %>%
    filter(pub_id == i) %>%
    dplyr::select(pub_id, days_enrollment, cens, vl, log10vl)
  # for error checking
  virus_rows = nrow(vl_data_ptid)
  
  obs_data = subset(indiv_preds, pub_id == i) %>%
    dplyr::select(pub_id, days_enrollment = time, obs_conc = DV) %>%
    mutate(days_enrollment = round(days_enrollment))
  obs_rows = nrow(obs_data)
  
  out = sim_pk(min(ptid_time_demo$art_day, ptid_time_demo$final_day), parms = parms, dose_info = dosing) %>%
    mutate(pub_id = i) %>%
    dplyr::select(pub_id, days_enrollment, centr_conc = Cc, peri_conc = Pc) %>%
    left_join(pk_indiv_parms, by = "pub_id") %>%
    left_join(dose_info_merge, by = c("pub_id", "days_enrollment")) %>%
    mutate(recent_infusion_time = na.locf(recent_infusion_time),
           infusion_no = na.locf(infusion_no)) %>%
    left_join(ptid_time_demo, by = "pub_id") %>%
    full_join(obs_data, by = c("pub_id", "days_enrollment")) %>%
    full_join(vl_data_ptid, by = c("pub_id", "days_enrollment"))

  # these just confirm the full joins worked correctly  
  stopifnot(all(!is.na(out$centr_conc)))
  stopifnot(all(!is.na(out$peri_conc)))
  stopifnot(nrow(subset(out, !is.na(log10vl))) == virus_rows)
  stopifnot(nrow(subset(out, !is.na(obs_conc))) == obs_rows)
  
  # check dose merge
  stopifnot(all(dosing$TIME %in% unique(out$recent_infusion_time)))
  stopifnot(all(dosing$DOSENO %in% unique(out$infusion_no)))
  
  out
}) %>%
  mutate(
    first_positive = days_enrollment == fp_day,
    diagnosis = days_enrollment == dx_day,
    post_fp = days_enrollment >= fp_day,
    vl_detected = if_else(!is.na(vl), cens != 1, F)
  )


stopifnot(nrow(subset(sim_pk_dat, diagnosis)) == 98)
n_distinct(subset(sim_pk_dat, diagnosis & vl_detected)$pub_id)

stopifnot(nrow(subset(sim_pk_dat, first_positive)) == 98)
n_distinct(subset(sim_pk_dat, first_positive & vl_detected)$pub_id)


n_distinct(subset(sim_pk_dat, vl_model_flag)$pub_id)

```


```{r error-checks}

# check matches sims up to first positive
stopifnot(all.equal(select(subset(sim_pk_dat, first_positive), names(sim_pk_prefp)),
          subset(sim_pk_prefp, first_positive), 
          tolerance = 1e-7)
          )

```

## PK data with vl measurement times

Note that from 52 people recieved infusions at/after dx and 60 at/after first positive

```{r}
#| fig.width=12, fig.height=12

sim_pk_dat %>%
  dplyr::filter(final_infusion_day < fp_day & days_enrollment >= final_infusion_day) %>%
  mutate(vl_days = if_else(vl_detected, days_enrollment, NA_real_)) %>%
  ggplot(aes(x = days_enrollment / 7 - fp_day / 7, y = pmax(0.1, centr_conc))) +
  geom_line() +
  scale_y_log10() +
  geom_vline(aes(xintercept = art_day / 7 - fp_day / 7), colour = "red") +
  geom_vline(aes(xintercept = vl_days / 7 - fp_day / 7), linetype = "dashed") +
  facet_wrap(~ pub_id, scales = "free_x") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "Wks post-fp") +
  ggtitle("PTIDs with final infusions before first postive. dashed black = pos. vl, red = ART time")

```

```{r}
#| fig.width=12, fig.height=12

sim_pk_dat %>%
  dplyr::filter(final_infusion_day >= fp_day & post_fp) %>%
  mutate(vl_days = if_else(vl_detected, days_enrollment, NA_real_)) %>%
  ggplot(aes(x = days_enrollment / 7 - fp_day / 7, y = centr_conc)) +
  geom_line() +
  xlab("Wks post-fp") +
  scale_y_log10() +
  geom_vline(aes(xintercept = art_day / 7 - fp_day / 7), colour = "red") +
  geom_vline(aes(xintercept = vl_days / 7 - fp_day / 7), linetype = "dashed") +
  facet_wrap(~ pub_id, scales = "free_x") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "Wks post-fp") +
  ggtitle("PTIDs with infusions on/after first postive. dashed black = pos. vl, red = ART time")

```

When dose times and viral load measurements overlap, want to make sure the vl measurement
is tagged before the dosing.

```{r}

sim_pk_dat %>%
  dplyr::filter(first_positive) %>%
  ggplot(aes(x = log10(pmax(centr_conc, 1)), y = log10vl)) +
  geom_point() +
  geom_smooth(method = "lm",  formula = 'y ~ x') +
  xlim(0, 1.75) +
  ylim(0, 8)

```

Confirming that VL measurement is just before the dose event on first positive.

```{r}

sim_pk_dat %>%
  dplyr::filter(days_enrollment == fp_day + 1 | first_positive) %>%
  mutate(log10vl = zoo::na.locf(log10vl), fp_cat = days_enrollment - fp_day) %>%
  ggplot(aes(x = log10(pmax(centr_conc, 1)), y = log10vl)) +
  geom_point() +
  #xlim(0, 1.75) +
  #ylim(0, 8) +
  facet_wrap(~fp_cat, nrow = 2) +
  ggtitle("PTIDs with infusions on/after first positive (fp day = 0, day after fp = 1).")

```

Only one person multiple (2) infusions after first positive.

```{r}

sim_pk_dat %>%
  dplyr::filter(final_infusion_day >= fp_day & post_fp) %>%
  group_by(pub_id) %>%
  summarize(
    post_fp_doses = n_distinct(infusion_no)
  ) %>%
  filter(post_fp_doses > 2)

sim_pk_dat %>%
  dplyr::filter(final_infusion_day >= fp_day & post_fp) %>%
  group_by(pub_id) %>%
  mutate(
    post_fp_doses = n_distinct(infusion_no)
  ) %>%
  filter(post_fp_doses > 2) %>%
  dplyr::filter(days_enrollment == final_infusion_day + 1 | 
                  days_enrollment == final_infusion_day) %>%
  select(pub_id, days_enrollment, final_infusion_day, log10vl, centr_conc)

```

## Fits - Simulation

```{r fits-rxode}

#| fig.width=12, fig.height=12

pk_fits_ode = ggplot(sim_pk_dat, aes(x = days_enrollment/7, y = pmax(0.03, centr_conc))) +
  geom_line() +
  geom_point(aes(y = obs_conc)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 10, ncol = 10) +
  geom_vline(aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks post-enrollment") +
  ggtitle("red = first positive")

pk_fits_ode

ggsave(here("output/pk-modeling/pk_fits_ode.pdf"), pk_fits_ode, width = 12, height = 12)

```

## Save key output

Merge on neut data for full time series PKPD with viral load.

Also filter to infection data alone.

```{r save-pk-sims}
glimpse(sim_pk_dat)
write_csv(sim_pk_dat, here("output/pk-modeling/sim-pk-withVLneut.csv"))

```

```{r save-shared-pk, eval = save_shared_folder}
write_csv(tibble(), 
           paste0("/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/!PK_SIM_DATA UPDATED - ",
                  Sys.Date()))

write_csv(sim_pk_dat, "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/sim-pk-withVLneut.csv")
```
