---
title: "Fitting PK Model"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

Here, we just use Lily's case-control PK model, "refit" the model to get the individual parameters and save the fitted parameters for each individual for future simulation work.
For participants without concentration data (there are 15), use the population-level parameters.

# Refit PK

```{r load-packages, echo = T, message=F,output=F}
knitr::opts_chunk$set(echo = T)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(lixoftConnectors)
library(RxODE)

theme_set(theme_bw() + theme(legend.position = "top"))
options(scipen = 999)
initializeLixoftConnectors(software = "monolix")

```

```{r load-funs-data}

source(here("R", "directory-funs.R"))
source(here("R", "mlx-model-funs.R"))
source(here("R", "fix-parm-funs.R"))
source(here("R", "mlx-res-processing.R"))

rx_dat = read_csv(raw_data_here("rx_dat.csv"), col_types = cols())
pk_data_nm = read_csv(mlx_data_here("pk-nm-data.csv"), col_types = cols())
median_weight = pk_data_nm %>% distinct(pub_id, weight) %>% summarize(median(weight)) %>% pull()
pk_data_analysis = read_csv(clean_data_here("adata-pk.csv"), col_types = cols())

# the following are for error checking
time_data = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols())
analysis_ptids = unique(pk_data_analysis$pub_id)

pk_dir_tag = function(x) paste0("PK/", x)
pk_model_here = function(x, ...) mlx_model_here(pk_dir_tag(x), ...)


rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

```

```{r refit-pk, eval = F}

# check_data("pk-nm-data.csv")

pk_data_types = tribble(
    ~var_name, ~type,
    "ID", "ignore",
    "AMT", "amount",
    "DV", "observation",
    "RATE", "rate",
    "TIME", "time",
    "dose", "ignore",
    "AVISITN", "ignore",
    "DOSENO", "ignore",
    "weight", "contcov",
    "study", "catcov",
    "pub_id", "id"
  )

amp_recipe(
  model_text_file = "infusion_2cpt_ClV1QV2.txt",
  data_file = "pk-nm-data.csv",
  variable_types = pk_data_types,
  initial_estimates = FALSE,
  error_model = "proportional",
  obs_dist = "normal"
)

fix_pk_parms_cc()

runScenario()
save_project(pk_dir_tag("VRC01AMP_Case-Cntrl"))
pop_ests = get_pop_ests(pk_dir_tag("VRC01AMP_Case-Cntrl"))
pop_ests %>% # this fixes a bug related to gpmx if i want to use it
  mutate(
    se_sa = NA_real_,
    rse_sa = NA_real_
    ) %>%
 write_csv(pk_model_here("VRC01AMP_Case-Cntrl/populationParameters.txt"))

```

# Processing/Error checking

Need to use pop ests for ptids without concentration data

```{r create-indiv-data}

# getIndividualParameterModel()$formula
# study = 0,1, tWT = log(weight/67.3)
# log(Cl) = log(Cl_pop) + beta_Cl_tWT*tWT + beta_Cl_study_1*[study = 1]
# log(V2) = log(V2_pop) + beta_V2_study_1*[study = 1]

pop_ests = get_pop_ests(pk_dir_tag("VRC01AMP_Case-Cntrl"))

pop_wide = spread(dplyr::select(pop_ests, -se_sa, -rse_sa),  key = parameter, value = value, drop = T)

indiv_parms_no_followup = pk_data_analysis %>%
  dplyr::filter(no_followup) %>%
  transmute(
    pub_id = pub_id,
    Cl = exp(log(pop_wide$Cl_pop) + pop_wide$beta_Cl_tWT*log(weight/67.3) + pop_wide$beta_Cl_study_1*(study == 1)),
    V1 = pop_wide$V1_pop,
    Q = pop_wide$Q_pop,
    V2 = exp(log(pop_wide$V2_pop) + pop_wide$beta_V2_study_1 * (study == 1)),
    est_method = 'population - missing followup' 
  )
stopifnot(nrow(indiv_parms_no_followup) == 15)

indiv_parms_mode = get_indiv_parms(pk_dir_tag("VRC01AMP_Case-Cntrl")) %>%
  dplyr::select(pub_id = id, contains("_mode")) %>%
  rename_with(~str_replace(., "_mode", "")) %>%
  dplyr::filter(pub_id %in% analysis_ptids) %>%
  mutate(est_method = 'mode') %>%
  bind_rows(indiv_parms_no_followup) %>%
  left_join(rx_dat, by = "pub_id") %>%
  mutate(rx_lab = factor(rx_code2, levels = rx_lab$rx_code2, labels = rx_lab$trt_cat)) %>%
  dplyr::select(-rx)

stopifnot(n_distinct(indiv_parms_mode$pub_id) == 96)
stopifnot(all(!is.na(indiv_parms_mode)))

missing_ptids_parms = setdiff(analysis_ptids, unique(indiv_parms_mode$pub_id))
stopifnot(n_distinct(missing_ptids_parms) == 0)

# the 15 people with no concentration data
missing_ptids_fits = setdiff(analysis_ptids, subset(indiv_parms_mode, est_method == "mode")$pub_id)
stopifnot(n_distinct(missing_ptids_fits) == 15)

```

# Save parameters

```{r save}

write_csv(indiv_parms_mode, here("output/pk-modeling", "pk-mode-parms.csv"))

```
