---
title: "Simulate PK Data"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

PK parameters fit based on case-control data (all infusions prior to first positive RNA).

Two simulations performed here:

1. "Case-Control": PK through first positive
2. "AMP PK": PK through end of study

Both datasets will have key times merged on for easier filter to concentrations at those times.

The dosing data comes from the full infusion information rather than the PK dataset, but results from both approaches are compared.
For the full infusion data, any infusion after day 0, is assumed to occur at time x.1 to offset with viral measurements that would be pre-infusion on the same day (days_enrollment_dose, see final step in data-processing).

```{r load-packages, echo = FALSE, message=F,output=F}
knitr::opts_chunk$set(echo = F)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(RxODE)

theme_set(theme_bw() + theme(legend.position = "top"))

options(scipen = 999)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

save_shared_folder = F

```


```{r load-funs-data}

source(here("R", "directory-funs.R"))
source(here("R", "mlx-res-processing.R"))

#rx_dat = read_csv(raw_data_here("rx_dat.csv"), col_types = cols())
pk_data_nm = read_csv(mlx_data_here("pk-nm-data.csv"), col_types = cols())
pk_data_analysis = read_csv(clean_data_here("adata-pk.csv"), col_types = cols())
time_data = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols())
all_dose_data = read_csv(clean_data_here("dosing-data.csv"), col_types = cols())
pk_indiv_parms = read_csv(here("output/pk-modeling", "pk-mode-parms.csv"), col_types = cols())

analysis_ptids = unique(pk_data_analysis$pub_id)
stopifnot(length(analysis_ptids) == 98)

vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols())

pk_location = "PK/VRC01AMP_Case-Cntrl"
pk_model_here = function(x, ...) mlx_model_here(pk_dir_tag(x), ...)
```


```{r pk-rxode-setup}

pk_2cmpt <- RxODE({
  kel = Cl/V1
  k12 = Q/V1
  k21 = Q/V2

  d/dt(centr) = -(k12+kel)*centr + k21 * peri
  d/dt(peri) = k12*centr - k21*peri
  
  Cc = centr/V1
  Pc = peri/V2
})

#' RxODE PK Simulator - simulated individual PK
#'
#' @param final_day 
#' @param parms tibble or named vector, needs Cl, V1, V2, Q
#' @param dose_info tibble/data.frame with dose info by columns for TIME, AMT, RATE, DOSENO
#' @param key_times vector of specific times to check 
#' @param post_infusion_draw automatically adds draw after infusions for correct peak (default 0 days)
#' @param output_grid_by time grid spacing (default daily)
#'
#' @return
#'
#' @examples 
#' test_sim = sim_pk(100, parms = tibble(Cl = 0.58, V1 = 3.53, Q = 0.676, V2 = 4.75),
#'   dose_info = tibble(TIME = c(0, 49.1), AMT = c(580, 600), RATE = c(15000, 27000), DOSENO = 1:2))
#' ggplot(test_sim, aes(x = days_enrollment, y = centr)) + geom_line()
#' 
sim_pk <- function(final_day,
                   parms,
                   dose_info,
                   key_times = NULL,
                   post_infusion_draw = 0,
                   output_grid_by = 1) {
  
  # add samples right after infusions
  post_dose_times = dose_info$TIME + dose_info$AMT/dose_info$RATE + post_infusion_draw
  
  sim_times = c(seq(0, final_day, by = output_grid_by), key_times, dose_info$TIME, post_dose_times) %>%
    sort() %>%
    unique()

  ev = eventTable(time.units = "days") %>%
    add.sampling(sim_times) 
  
  for(i in dose_info$DOSENO){
    dose_iteration = dplyr::filter(dose_info, DOSENO == i)
    ev = ev %>%
      add.dosing(
      dose = dose_iteration$AMT,
      rate = dose_iteration$RATE,
      nbr.doses = 1,
      cmt = "centr",
      start.time = dose_iteration$TIME
      )
  }

  as_tibble(rxSolve(pk_2cmpt, params = parms, event = ev)) %>% 
    mutate(days_enrollment = as.numeric(time)) %>%
    dplyr::select(-time)

}

```

# Explore PK

```{r read-pk-res, eval = F}
#loadProject(mlx_model_here(pk_dir_tag("VRC01AMP_Case-Cntrl"), mlxtran = T))
indiv_fits = get_indiv_fits(pk_location) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids) 

indiv_preds = get_indiv_preds(pk_location) %>% rename(pub_id = ID) %>%
  dplyr::filter(pub_id %in% analysis_ptids)
```

## Fits - mlx

```{r, fits}
#| fig.width=12, fig.height=12

pk_fits = ggplot(indiv_fits, aes(x = time/7, y = pmax(0.03, indivPredMode))) +
  geom_line() +
  geom_point(data = indiv_preds, aes(y = DV)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 10, ncol = 9) +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = tpf_day/7), colour = "red", linetype = 'dashed') +
  geom_vline(data = dplyr::filter(time_data, pub_id %in% unique(indiv_fits$pub_id)), 
             aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks") +
  ggtitle("dashed = PF-infection time. solid = first positive")

pk_fits

ggsave(here("output/pk-modeling/pk_fits.pdf"), pk_fits, width = 12, height = 12)

```

# Sims - RxODE

## Dose error checking

Two participants have infusions that were not in the case-control data.

The times are rounded in my cleaned version of the dose data. RATES are calculated differently, when done with SAS
there was rounding before unit conversion (differences < 1 minute).

```{r setup-doses}

dose_data_casectrl = pk_data_nm %>%
  dplyr::filter(DV == "." & AMT != "." & pub_id %in% analysis_ptids) %>%
  mutate(across(c(AMT, RATE), as.numeric)) %>%
  select(pub_id, TIME, AMT, RATE, DOSENO) %>%
  arrange(pub_id, TIME, AMT, RATE, DOSENO)

dose_data = all_dose_data %>%
  dplyr::filter(pub_id %in% analysis_ptids) %>%
  select(pub_id, TIME = days_enrollment_dose, AMT = dose_mg, RATE, DOSENO = infusion_no, days_fp) %>%
  arrange(pub_id, TIME, AMT, RATE, DOSENO)
  
dose_data_prefp = dplyr::filter(dose_data, days_fp < 0)

# for error we just compare the pre-first positive to case-control

# should be missing .x not .y
merge_test = full_join(dose_data_casectrl, dose_data_prefp, by = c("pub_id", "DOSENO"))
stopifnot(all(!is.na(merge_test$AMT.y)))
subset(merge_test, is.na(AMT.x))

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = AMT.y-AMT.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = TIME.y-TIME.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

ggplot(subset(merge_test, !is.na(AMT.x)), aes(x = pub_id, y = RATE.y/RATE.x)) +
  geom_point() +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())

# reduce_comp = diffdf::diffdf(base = dose_data_casectrl, 
#                              subset(select(dose_data_prefp,  -days_fp), !( (pub_id == "704-2624" & DOSENO == 9) |
#                                                           (pub_id == "704-2901" & DOSENO == 8))))

```

## Simulation A: Through first positive

WHY ARE THERE NEGATIVE ENROLLMENT DAYS!?

```{r sim-pk-fp}

sim_pk_data_prefp = map_df(unique(pk_indiv_parms$pub_id), function(i){
  
  parms = dplyr::select(subset(pk_indiv_parms, pub_id == i), Cl, V1, Q, V2)
  stopifnot(nrow(parms) == 1)
  
  dosing = subset(dose_data_prefp, pub_id == i) 
  dosing_cc = subset(dose_data_casectrl, pub_id == i) 
  
  fp_day = subset(time_data, pub_id == i)$fp_day
  
  
  vl_data_prefp = vl_data %>%
    filter(pub_id == i & days_enrollment <= fp_day) %>%
    dplyr::select(pub_id, days_enrollment, cens, vl, log10vl)
  
  test_sim = sim_pk(fp_day, parms = parms, dose_info = dosing_cc) %>%
    mutate(temp = "case-control-dosing", pub_id = i)
       
  out = sim_pk(fp_day, parms = parms, dose_info = dosing) %>%
    mutate(temp = "dosing-data", pub_id = i) %>%
    bind_rows(test_sim) %>%
    dplyr::select(pub_id, days_enrollment, centr_conc = Cc, peri_conc = Pc, temp) %>%
    left_join(pk_indiv_parms, by = "pub_id") %>%
    full_join(vl_data_prefp, by = c("pub_id", "days_enrollment")) %>%
    mutate(fp_day =  fp_day)
    
  
  if(any(is.na(out$centr_conc))) browser()
  
  stopifnot(all(!is.na(out$centr_conc)))
  stopifnot(all(!is.na(out$peri_conc)))
  
  out
})

stopifnot(nrow(subset(sim_pk_data_prefp, temp == "dosing-data" & days_enrollment == fp_day)) == 98)

```

```{r comp-fp-conc}

sim_pk_data_prefp %>%
  dplyr::filter(days_enrollment == fp_day) %>%
  select(pub_id, temp, centr_conc) %>%
  pivot_wider(values_from = centr_conc, names_from = temp) %>%
  filter(abs(`dosing-data` - `case-control-dosing`) > 0.1) %>%
  ggplot(aes(x = pub_id, y = `dosing-data` -`case-control-dosing`)) +
    geom_point()

```


```{r old}


sim_pk_dat = map_df(unique(subset(indiv_parms_mode, !is.na(Cl))$pub_id), function(i){
  
  # relevent data
  parms = dplyr::select(subset(indiv_parms_mode, pub_id == i), Cl, V1, Q, V2)
  dosing = subset(dose_data, pub_id == i) 
  vl_data = dplyr::select(subset(vl_data_analysis, pub_id == i & days_enrollment >= fp_day), 
                          pub_id, days_enrollment, cens, vl, log10vl)
  vl_data_times = vl_data$days_enrollment
  vl_times = subset(time_data, pub_id == i)
  obs_data = dplyr::select(subset(indiv_preds, pub_id == i), pub_id, time, obs_conc = DV)
  
  final_time = vl_times$final_vl_day
  extra_pk_vl_times = c(obs_data$time,
               subset(indiv_fits, pub_id == i)$time,
               vl_data_times, 
               vl_times$tpf_day
  )
               
  sim_pk(
    vl_times$final_vl_day,
    parms = parms,
    dose_info = dosing,
    key_times = extra_pk_vl_times
  ) %>%
    mutate(pub_id = i) %>%
    dplyr::select(pub_id, days_enrollment, centr_conc = Cc, peri_conc = Pc) %>%
    left_join(indiv_parms_mode, by = "pub_id") %>%
    left_join(obs_data, by = c("pub_id", "days_enrollment" = "time")) %>%
    left_join(vl_data, by = c("pub_id", "days_enrollment")) %>%
    mutate(
      tpf_day = vl_times$tpf_day,
      fp_day =  vl_times$fp_day,
      dx_day = vl_times$dx_day,
      post_fp = days_enrollment >= fp_day,
      vl_detected = !is.na(vl),
    ) 
  
})

```

```{r fits-rxode}

#| fig.width=12, fig.height=12

pk_fits_ode = ggplot(sim_pk_dat, aes(x = days_enrollment/7, y = pmax(0.03, centr_conc))) +
  geom_line() +
  geom_point(aes(y = obs_conc)) +
  scale_y_log10() +
  facet_wrap( ~ pub_id, nrow = 10, ncol = 10) +
  geom_vline(aes(xintercept = tpf_day/7), colour = "red", linetype = 'dashed') +
  geom_vline(aes(xintercept = fp_day/7), colour = "red") +
  labs(y = "[VRC01]") +
  theme(legend.position = 'top') +
  labs(x = "weeks") +
  ggtitle("dashed = PF-infection time. solid = first positive")

pk_fits_ode

ggsave(here("output/pk-modeling/pk_fits_ode.pdf"), pk_fits_ode, width = 12, height = 12)

```

## Save key output

Merge on neut data for full time series PKPD with viral load.

Also filter to infection data alone.

```{r save-pk-sims}

sim_pk_dat_neut = sim_pk_dat %>%
  left_join(dplyr::select(neut_data, -contains("cat")) , by = "pub_id") %>%
  write_csv(here("output/pk-modeling/pk-sims-vl-neut.csv"))

infection_day_conc = sim_pk_dat_neut %>%
  dplyr::filter(abs(days_enrollment - tpf_day) <1e-6) %>%
  dplyr::select(pub_id, days_enrollment, centr_conc, peri_conc, tpf_day, fp_day, contains("rx"), contains("gmt")) %>%
   write_csv(here("output/pk-modeling/infection-day-pkpd.csv"))

n_distinct(infection_day_conc)
infection_day_conc %>% group_by(pub_id) %>% summarize(total = n()) %>% filter(total != 1)

```


```{r}
read_csv(here("output/pk-modeling/infection-day-pkpd-x.csv")) %>%
  dplyr::select(pub_id, centr_conc) %>%
  full_join(select(infection_day_conc, pub_id, centr_conc), by = "pub_id") ->x

```

```{r save-shared-pk, eval = save_shared_folder}
write_csv(sim_pk_dat_neut, "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/sim-pk-withVLPD.csv")
write_csv(infection_day_conc, "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/predicted-pk-infection-day.csv")
```


```{r test-dat-setup, eval = F}

# this just tests consistency between simulations
# the alignment of continuous sample times is important for concentration

test_pub_id = "704-0662"
test_set = subset(pk_indiv_parms, pub_id == test_pub_id) %>%
  select(where(is.numeric))
test_dose = subset(dose_data, pub_id == test_pub_id) 
test_vl_times = subset(time_data, pub_id == test_pub_id)

x = sim_pk(test_vl_times$fp_day, test_set, dose_info = test_dose, 
           key_times = c(round(subset(indiv_preds, pub_id == test_pub_id)$time, 1),
                         test_vl_times$fp_day)) %>%
  mutate(pub_id = test_pub_id)


ggplot(data = x, aes(x = days_enrollment/7, y = pmax(0.1,Cc))) +
  geom_line() +
  geom_line(data = subset(indiv_fits, pub_id == test_pub_id), 
            aes(x = time/7, y = pmax(0.1, indivPredMode)), colour = "red") +
  scale_y_log10(limits = c(0.09, NA)) +
  geom_vline(data = test_vl_times, aes(xintercept = fp_day/7)) 

subset(indiv_preds, pub_id == test_pub_id) %>%
  mutate(time = round(time, 1)) %>%
  dplyr::select(pub_id, time, DV, indivPredMode) %>%
  left_join(dplyr::select(mutate(indiv_fits, time = round(time, 1)), pub_id, time, indivPredMode), by = c("pub_id", "time")) %>%
  left_join(dplyr::select(x, pub_id, time = days_enrollment, Cc), by = c("pub_id", "time")) 

```
