---
title: "Cohort tables for VL analysis"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
---

```{r load-packages, echo = FALSE, message=F,output=F,warning=F}
knitr::opts_chunk$set(echo = F)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(cowplot)
library(janitor)
library(table1)
library(flextable)
```


```{r setup}
theme_set(theme_bw() + theme(legend.position = "top", panel.grid.minor = element_blank()))

source(here("R", "directory-funs.R"))

paste_range0 = function(x) paste_range(x, digits = 0)
paste_range = function(x, digits = 2) paste(round(range(x), digits = digits), collapse = ", ")

#stopifnot(dir.exists("/Volumes/trials/"))

vl_lims = c(1, 7)
ic_cats = c("< 1 mcg/mL", "1-3 mcg/mL", "> 3 mcg/mL")
vl_title = expression(paste("Viral load (log"[10], " copies/mL)"))
ep_labels = tibble(
  levels = c("fp", "peak", "nadir", "mean"),
  label = c("First positive", "Peak", "Nadir", "Mean")
)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

```


```{r load-data}

measurement_summary = read_csv(clean_data_here("ptid-measurement-counts.csv"), col_types = cols())
vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols())
time_summary = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols())
dose_data = read_csv(clean_data_here("dosing-data.csv"), col_types = cols())

full_vl_data = read_csv(clean_data_here("full-vl-data.csv"), col_types = cols())

rx_dat = read_csv(raw_data_here("rx_dat.csv"), col_types = cols())

rx_map = rx_dat %>% select(pub_id, rx_code2) %>%
  mutate(rx_lab = factor(rx_code2, levels = rx_lab$rx_code2, labels = rx_lab$trt_cat),
         .keep = "unused")

study_region_map = distinct(vl_data, study_region_code, study_region_cat) %>%
  mutate(study_region_cat2 = if_else(str_detect(study_region_cat, "703"), 
                                     paste0(study_region_cat, "/081"),
                                     paste0(study_region_cat, "/085")))

demo_vars = distinct(vl_data, pub_id, study_region_code, any_art, protocol) %>%
  left_join(rx_map, by = "pub_id") %>%
  mutate(study_region_cat = factor(study_region_code, 
                                   levels = study_region_map$study_region_code,
                                   labels = study_region_map$study_region_cat2),
         any_art = factor(any_art, levels = c(F, T), labels = c("No", "Yes")))

neut_vars = distinct(vl_data, pub_id, nisolates, gmt80ls) %>%
  mutate(ic80gte1 = gmt80ls>=1)

t1_print_save = function(t1, file){
  save_as_docx(t1flex(t1), path=paste0("tables/", file, ".docx"))
  t1
}

```

# Background

Outline the cohort information, and highlight data that we need as we are performing the analysis. 

  - Sample sizes and total measurements, designated by pre- and post-ART
  - By group and protocol
  - Baseline covariates:
    - Geo-location/protocol (AMP paper): (1) South Africa/703, (2) not South Africa/703, (3) US or Switzerland/704, (4) South America/704.  
  - Other covariates:
    - total isolates
    - neutralization (how do we downselect this variable). least sensitive IC80 used in NEJM paper
  
PG: I would think this table would include information on available VL data pre and post ART initiation for all of the scheduled viral load sampling visits.  And indeed the basic flow stuff about numbers randomized, numbers acquiring HIV, and of those numbers with viral load data by visit. Ideally the table would orient to the data available by IC80 >= 1 and IC80 < 1


# Results

## Table 1s

```{r prep-table1}

table1_prep = measurement_summary %>%
  dplyr::filter(key %in% c("total_measurements", "total_measurements_offart")) %>%
  dplyr::select(pub_id, key, total) %>%
  pivot_wider(names_from = "key", values_from = "total") %>%
  dplyr::mutate(across(c(total_measurements_offart, total_measurements), 
                       ~fct_relevel(factor(if_else(. < 5, as.character(.), "> 4")),
                                    "> 4", after = Inf))
  ) %>%
  left_join(demo_vars, by = "pub_id") %>%
  left_join(neut_vars, by = "pub_id") %>%
  mutate(
    ic80gte1 = factor(ic80gte1, levels = c(F, T), 
                      labels = c("< 1 mcg/mL", bquote(">= 1 mcg/mL"))),
    nisolates = factor(nisolates)
    ) 

label(table1_prep$study_region_cat) = "Region/Protocol"
label(table1_prep$ic80gte1) = "IC80 (least sensitive)"
label(table1_prep$nisolates) = "Total isolates"
label(table1_prep$any_art) = "ART use"
label(table1_prep$total_measurements) = "Total VL measurements <br>(post-first positive)"
label(table1_prep$total_measurements_offart) = "Total VL measurements<br>(pre-ART)"


var_formula = "~ study_region_cat + any_art + ic80gte1 + nisolates + total_measurements + total_measurements_offart"

```

### By VRC01 groups

```{r table1-grp-overall}

table1(as.formula(paste(var_formula, "| rx_lab")), 
       data=table1_prep, caption = "Demographic and Virologic Characteristics of AMP Participants with HIV-acquisition by VRC01 Group.")  %>%
  t1_print_save("main-table1")

```

### By protocol and VRC01 groups

```{r table1-bygroup}

table1(as.formula(paste(var_formula, "| rx_lab*protocol")),
        data=table1_prep, 
        caption = "Demographic and Virologic Characteristics of AMP Participants with HIV-acquisition by VRC01 Group and Protocol.") %>%
  t1_print_save("supp-table1-by-protocol")

```

# Measurements by visit

Visit schedule (Appendix G from the protocol).

The strategy is to break this up into two segments. Segment one is all pre-dx measurements. Segment two is post-dx. Listed by scheduled visit time and actual range. Want to look at on/off art together

```{r scheduled-visit}

dx_visit_schedule = tribble(
  ~visitno, ~expected_day, ~expected_wk,
  "v.x", "diagnosis", "W0",
  "v.y", "D14", "W2",
  "31", "D28", "W4",
  "32", "D42", "W6",
  "33", "D56", "W8",
  "34", "D84", "W12",
  "35", "D168", "W24"
)

dx_visit_schedule %>%
  write_csv("tables/supp-visitno_key.csv") %>%
  kable(caption = "visit no. by diagnosis.") %>%
  kable_styling(full_width = F) 

```


```{r pre-dx-fp}

fp_pre_dx = full_vl_data %>%
  dplyr::filter(days_dx < 0 & days_fp >= 0 &  infection_pre85wks & has_neut_data) %>%
  mutate(visitno_tmp = "Pre-Diagnosis")

fp_pre_dx_summary = fp_pre_dx %>%
  group_by(visitno_tmp, on_art) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_daydx_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx)),
    avg_dayfp_range = stat_paste(as.character(round(mean(days_fp), 1)), min(days_fp), max(days_fp)),
    .groups = "drop"
  ) 

fp_pre_dx_summary

```

Using the visits as coded. Need to resolve two issues, the dx day coding and the "Others". When recoding the others, maybe do categories like
"Off-schedule: WX-WY"

```{r dx-data}

dx_post_data = full_vl_data %>%
  filter(days_dx >= 0  &  infection_pre85wks & has_neut_data) %>%
  mutate(
    visitno_cat = case_when(
      visitno == dxvisit ~ "Dx visit",
      floor(visitno) == floor(dxvisit) &
      visitno != dxvisit ~ "First follow-up",
      !visitno %in% 31:35 ~ "Other",
      TRUE ~ as.character(visitno)
      )
  )

dx_post_summary = dx_post_data %>%
  group_by(visitno_cat, on_art) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_day_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx)),
    .groups = "drop"
  ) 

dx_post_summary %>%
  arrange(on_art, visitno_cat)

dx_post_summary %>%
  filter(on_art == 0) %>%  
  janitor::adorn_totals()

```

There are some visitno labeled for dx-day, but should probably actually be labeled with the 30 conventions.

```{r}

dx_day_dat = full_vl_data %>%
  dplyr::filter(has_neut_data & floor(visitno) == floor(dxvisit) & days_dx >= 0) %>%
   mutate(
     visitno_tmp_x = round(visitno - floor(dxvisit), 1),
     visitno_tmp_check = if_else(days_dx == 0, 0, round(visitno - floor(dxvisit), 1)),
     visitno_tmp = if_else(days_dx == 0, "v.x", "v.y")
     )

dx_visits_afterd22 = subset(dx_day_dat, days_dx>22) 

tmp = subset(vl_data, pub_id %in% dx_visits_afterd22$pub_id & days_dx > 0)
check_dx = ggplot(tmp, aes(x = visitno, y = days_dx)) + geom_point() + geom_line()+ facet_wrap(~pub_id)

dx_day_dat %>%
  group_by(visitno_tmp_x, on_art) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_day_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx))
  ) %>%
  arrange(on_art, visitno_tmp_x)

dx_day_dat %>%
  group_by(visitno_tmp_check) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_day_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx))
  ) 

dx_dat_summary = dx_day_dat %>%
  group_by(visitno_tmp, on_art) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_daydx_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx)),
    avg_dayfp_range = stat_paste(as.character(round(mean(days_fp), 1)), min(days_fp), max(days_fp)),
    .groups = "drop"
    ) 

dx_dat_summary

```

```{r post-dx-data}

post_dx_dat = vl_data %>%
  dplyr::filter(days_dx >= 0 & floor(visitno) != floor(dxvisit) & has_neut_data) %>%
  mutate(visitno_tmp = if_else(visitno %in% 31:35, as.character(floor(visitno)), "Off-schedule"))

post_dx_dat_summary = post_dx_dat %>%
  group_by(visitno_tmp) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_daydx_range = stat_paste(as.character(round(mean(days_dx), 1)), min(days_dx), max(days_dx)),
    avg_dayfp_range = stat_paste(as.character(round(mean(days_fp), 1)), min(days_fp), max(days_fp))
    ) 

```

```{r visitno-summary}

fp_pre_dx_summary %>%
  bind_rows(filter(dx_dat_summary, on_art == 0)) %>%
  bind_rows(post_dx_dat_summary) %>%
  ungroup() %>%
  rename(visitno = visitno_tmp) %>%
  left_join(dx_visit_schedule, by = "visitno") %>%
  select(visitno, N, n, `Expected Day` = expected_day, `Mean day post-dx (range)` = avg_daydx_range,
         `Mean day post-first pos (range)` = avg_dayfp_range) %>%
  adorn_totals() %>%
  write_csv("tables/supp-visit-summary.csv") %>%
  kable(caption = "VL Measurements by post-dx visit numbers and corresponding study days. Days post-diagnosis. N = total participants, n = total measurements.", digits = 2) %>%
  kable_styling(full_width = F)

```

```{r alt-cutoffs}

cuts = c(-125, 0, 1, 22, 38, 52, 70, 140, 197, 400) 

vl_data_dx_cut = vl_data %>%
  dplyr::filter(has_neut_data, days_fp >= 0) %>%
  mutate(
    days_dx_cut = cut(days_dx, breaks = cuts, include.lowest = T, right = F),
    ic80gte1 = factor(gmt80ls >= 1, levels = c(F, T), 
                      labels = c("IC80 < 1 mcg/mL", bquote("IC80 >= 1 mcg/mL"))),
    pre_art = factor(if_else(on_art == 0, "Pre-ART", "Post-ART"), levels = c("Pre-ART", "Post-ART")),
    nisolates = factor(nisolates)
    ) 
  
vl_data_dx_cut %>%
  group_by(days_dx_cut) %>%
  summarize(N = n_distinct(pub_id),
            n = n()) %>%
  dplyr::select(`Post-dx days range` = days_dx_cut, everything()) %>%
  adorn_totals() %>%
  write_csv("tables/other/measurement-days.csv") %>%
  kable(caption = "VL Measurements based on diagnosis date not visit. N = total participants, n = total measurements.", digits = 2) %>%
  kable_styling(full_width = F)

```

```{r, full-visit-setup}

counts_only = function (x, ..., na.is.category = TRUE) {
    c("", sapply(stats.apply.rounding(stats.default(x, ...), 
        ...), function(y) with(y, sprintf("%s", FREQ))))
}
label(vl_data_dx_cut$days_dx_cut) = "Post-dx study days range"

```


```{r by-art}

table1(~days_dx_cut|rx_lab, data = vl_data_dx_cut, render.categorical = counts_only, caption = "Measurement counts by VRC01 groups.") %>%
  t1_print_save("supp-measurement-counts-by-group")

```



```{r by-grp-ic80}

table1(~days_dx_cut|rx_lab + ic80gte1, data = vl_data_dx_cut, render.categorical = counts_only,
       caption = "Pre-ART measurement counts by VRC01 groups and IC80 cutoffs.") %>%
  t1_print_save("supp-measurement-counts-by-group-ic80")

```


```{r full-visitno-post-dx}

post_dx_dat %>%
  group_by(visitno) %>%
  summarize(
    N = n_distinct(pub_id),
    n = n(),
    avg_day_dx = mean(days_dx),
    day_dx_range = paste_range0(days_dx)
  ) %>%
  write_csv("tables/other/visitno-all.csv") %>%
  kable(caption = "Full description of post-dx visit numbers", digits = 2) %>%
  kable_styling(full_width = F)

```

## Demographics (more details)

```{r check-country}

vl_data %>%
  dplyr::filter(has_neut_data) %>%
  unite(protocol_country, protocol, country) %>%
  distinct(pub_id, study_region_cat, protocol_country) %>%
  group_by(study_region_cat, protocol_country) %>%
  tally() %>%
  spread(study_region_cat, n, fill = 0) %>%
  separate(protocol_country, into = c("protocol", "country"), sep = "_") %>%
  janitor::adorn_totals("col") %>%
  janitor::adorn_totals() %>%
  write_csv("tables/other/region-sample-size.csv") %>%
  kable(caption = "Study region breakdown among VL participants") %>%
  kable_styling(full_width = F)

```


# Infusion timing

```{r quick-tally-test}

# just confirms use of time_summary data, consistent with the dose data
# Dose data can easily tally how many doses after first positive though

test = time_summary %>%
  filter(rx_code2 != "C") %>%
  summarize(
    N = n(),
    total_infusions_at_after_dx = sum(final_infusion_day >= dx_day),
    total_infusions_at_after_fp =  sum(final_infusion_day >= fp_day)
  ) 

stopifnot(test$total_infusions_at_after_dx == 
            summarize(filter(dose_data, post_dx), n =  n_distinct(pub_id))$n)
stopifnot(test$total_infusions_at_after_fp == 
            summarize(filter(dose_data, post_fp), n =  n_distinct(pub_id))$n)

dose_data %>%
  filter(post_fp) %>%
  left_join(rx_map, by = "pub_id") %>%
  left_join(neut_vars, by = "pub_id") %>%
  group_by(pub_id, rx_lab, ic80gte1, gmt80ls) %>%
  summarize(
    n = n(),
    .groups = "drop"
  ) %>%
  filter(n > 1) 

```


```{r}
time_summary %>%
  group_by(group = rx_code2) %>%
  summarize(
    N = n(),
    last_infusion_at_after_dx = sum(final_infusion_day >= dx_day),
    last_infusion_at_after_fp =  sum(final_infusion_day >= fp_day),
    .groups = "drop"
  ) %>%
  janitor::adorn_totals() %>%
  write_csv("tables/other/infusion-all-grps.csv") %>%
  kable() %>%
  kable_styling(full_width = F)

```


```{r}

time_summary %>%
  left_join(neut_vars, by = "pub_id") %>%
  mutate(rx_pool = factor(rx_code2 != "C", levels = c("FALSE", "TRUE"),
                          labels = c("Placebo", "VRC01 (Pooled)"))) %>%
  group_by(group = rx_pool, `IC80 >= 1` = ic80gte1) %>%
  summarize(
    N = n(),
    last_infusion_at_after_dx = sum(final_infusion_day >= dx_day),
    last_infusion_at_after_fp =  sum(final_infusion_day >= fp_day),
    .groups = "drop"
  ) %>%
  janitor::adorn_totals() %>%
  write_csv("tables/supp-infusion-grps.csv") %>%
  kable(caption = "Total participant counts by whether last infusion was given after first positive. 1 ptid (10mg/kg, IC80 = 5) had two infusions at/after first positive. Participants in placebo group received saline infusions.") %>%
  kable_styling(full_width = F)

```