---
title: "Fit RSIF Holte model to AMP data"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

Here, we fit the RV217 Holte models to the AMP participants. Except for the infection time parameters, all of the fixed population parameters are fixed from the RSIF paper. The population parameters are never refit but the variance/correlations terms using different strategies.

```{r load-packages, echo = T, message=F,output=F, warning=F}
knitr::opts_chunk$set(echo = T)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(GGally)
library(lixoftConnectors)
source(here("R", "directory-funs.R"))
source(here("R", "mlx-model-funs.R"))
source(here("R", "mlx-envir-res.R"))
source(here("R", "mlx-res-processing.R"))
source(here("R", "fix-parm-funs.R"))
source(here("R", "mlx-plot-funs.R"))

initializeLixoftConnectors(software = "monolix")
theme_set(theme_bw() + theme(legend.position = "top"))
options(scipen = 999)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)
vl_dir_tag = function(x) paste0("VL/", x)
vl_model_here = function(x, ...) mlx_model_here(vl_dir_tag(x), ...)

```

```{r load-funs-data}

rv217_fits = read_csv(mlx_data_here('rsif-timing-parms.csv'), col_types = cols())
dan_pop_fits = read_csv(here('output/vl-analysis/populationParameters.txt'), col_types = cols()) %>%
  select(parameter, value) %>% spread(parameter, value)

demo_merge = read_csv(clean_data_here("adata-vl.csv"), col_types = cols()) %>%
  dplyr::select(pub_id, contains("gmt"), nisolates, protocol, study, rx_code2) %>% 
  distinct() %>%
  left_join(rx_lab, by = "rx_code2")
stopifnot(nrow(demo_merge) == 158)

```


```{r vl-type-setup}
# check_data("amp-mlx-vl.csv")

vl_types = tribble(
    ~var_name, ~type,
    "pub_id", "id",
    "TIME", "ignore",
    "DV", "observation",
    "DV_log", "ignore",
    "fp_day", "ignore",
    "days_fp", "TIME",
    "tpf_est", "ignore",
    "cens", "cens",
    "limit", "limit",
    "protocol", "catcov",
    "study_region_cat", "catcov",
    "gmt50ms", "ignore",
    "gmt80ms", "ignore",
    "gmt50ls", "ignore",
    "gmt80ls", "ignore",
    "gmt50mf", "ignore", 
    "gmt80mf", "ignore",
    "nisolates", "ignore",
    "study", "ignore",
    "rx_code2", "ignore"
  ) 

log10vl_types = vl_types
log10vl_types[log10vl_types$var_name == "DV", ]$type = "ignore"
log10vl_types[log10vl_types$var_name == "DV_log", ]$type = "observation"


```

# Model fitting

model tag: "holte_vl"


The major models:
**All models refit infection time parameter and omega (for time).**

- RSIF, all fixed: holte_vl_corr
- RSIF, all fixed correlation = 0: holte_vl
- refit correlation: holte_vl_refitcorr
  - Dan's model is this with initial conditions for correlations = 0 (I verified similar infection time draws but did not reproduce the fitting algorithm for this model)
- refit omega: holte_vl_omega
- refit omega and correlation: holte_vl_refitomegacorr

Two sets of sub-models:

- A) Pooled fit: Fit using all participants (denoted _all)
- B) Placebo trained
  - 1) Fit using placebo only (denoted _placebo)
  - 2) Placebo model applied to vrc01 groups (denoted _vrc01)

The outcome variable is vl ~ log normal.

## Fixing all parameters (w/ correlations)

```{r rsif-fixed-all-corr, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

# getPopulationParameterInformation()
set_holte_parms(holte_parms = rv217_fits, est_corr_rsif = "FIXED")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )

runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_corr_all"))

```

```{r rsif-fixed-placebo-corr, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-placebo.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

# getPopulationParameterInformation()
set_holte_parms(holte_parms = rv217_fits, est_corr_rsif = "FIXED")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )

runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_corr_placebo"))

```

```{r rsif-fixed-vrc01-corr, eval = F}

placebo_rsif = get_pop_ests(vl_dir_tag("holte_vl_corr_placebo")) %>%
  select(parameter, value) %>% spread(parameter, value)
  
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-vrc01.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = placebo_rsif, est_corr = "FIXED")
setPopulationParameterInformation(
  a = list(initialValue = placebo_rsif$a, method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )

check = left_join(getPopulationParameterInformation(), 
                                  gather(placebo_rsif), by = c("name" = "key"))
check
stopifnot(nrow(subset(check, 
                        abs(initialValue - value) > 1e-4 &
                          name != "a" &
                          !is.na(value) )
                 ) == 0
  )
  
runScenario()

save_project(vl_dir_tag("holte_vl_corr_vrc01"))

```

## Fixing all parameters (no correlations)

Based on later diagnostics, this model has too tight variability between
participants.

```{r rsif-fixed-all, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

# getPopulationParameterInformation()
set_holte_parms(holte_parms = rv217_fits)

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )

runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_all"))

```

```{r rsif-fixed-placebo, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-placebo.csv",
  variable_types = vl_types,
  initial_estimates = FALSE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = rv217_fits)

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )
runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_placebo"))

```

## Refitting correlations

```{r rsif-fixed-all-refitcorr, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

# getPopulationParameterInformation()
set_holte_parms(holte_parms = rv217_fits, est_corr_rsif = "MLE")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )

runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_refitcorr_all"))

```

```{r rsif-fixed-placebo-refitcorr, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-placebo.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

# getPopulationParameterInformation()
set_holte_parms(holte_parms = rv217_fits, est_corr_rsif = "MLE")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )

runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_refitcorr_placebo"))

```


```{r rsif-refit-corr-trt, eval = F}

placebo_refitcorr = get_pop_ests(vl_dir_tag("holte_vl_refitcorr_placebo")) %>%
  select(parameter, value) %>% spread(parameter, value)
  
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-vrc01.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = placebo_refitcorr, est_corr = "FIXED")
setPopulationParameterInformation(
  a = list(initialValue = placebo_refitcorr$a, method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )

check = left_join(getPopulationParameterInformation(), 
                                  gather(placebo_refitcorr), by = c("name" = "key"))
check
stopifnot(nrow(subset(check, 
                        abs(initialValue - value) > 1e-4 &
                          name != "a" &
                          !is.na(value) )
                 ) == 0
  )
  
runScenario()

save_project(vl_dir_tag("holte_vl_refitcorr_vrc01"))

```

## Reftting omegas

```{r rsif-refit-omega-all, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = rv217_fits, est_omega = "MLE", est_corr_rsif = "FIXED")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )
runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_refitomega_all"))

```

```{r rsif-refit-omega-placebo, eval = F}
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-placebo.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = rv217_fits, est_omega = "MLE",  est_corr_rsif = "FIXED")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )
runScenario()
get_pop_parms()

save_project(vl_dir_tag("holte_vl_refitomega_placebo"))

```

```{r rsif-refit-omega-trt, eval = F}

placebo_refitomega = get_pop_ests(vl_dir_tag("holte_vl_refitomega_placebo")) %>%
  select(parameter, value) %>% spread(parameter, value)
  
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-vrc01.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = placebo_refitomega, est_corr_rsif = "FIXED")
setPopulationParameterInformation(
  a = list(initialValue = placebo_refitomega$a, method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )

check = left_join(getPopulationParameterInformation(), 
                                  gather(placebo_refitomega), by = c("name" = "key"))
check
stopifnot(nrow(subset(check, 
                        abs(initialValue - value) > 1e-4 &
                          name != "a" &
                          !is.na(value) )
                 ) == 0
  )
  
runScenario()

save_project(vl_dir_tag("holte_vl_refitomega_vrc01"))

```

## Refitting omega/correlations

```{r rsif-refit-omegacorr-all, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = rv217_fits, est_omega = "MLE", est_corr_rsif = "MLE")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )
runScenario()

save_project(vl_dir_tag("holte_vl_refitomegacorr_all"))

```

Started with the RV217 RSIF model with free omegas and correlation. Initial fits suggest the
dI correlations are minor and poorly specified (high RSEs) so I remove all 3 (they also switched direction from negative to positive). After fixing the constant error 
(because the se is always NaN) and refitting the omega dS was poorly specified (NaN, standard error). Then I tried linearization and that seemed to work well, none of the rses shrunk and omega dS se was estimated.

```{r rsif-refit-omegacorr-placebo, eval = F}

amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-placebo.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = rv217_fits, est_omega = "MLE", est_corr_rsif = "MLE")

# get initial estimates w/ simulated annealing on
runScenario()
setInitialEstimatesToLastEstimates()
set_project_settings(initial_estimates = FALSE)
runScenario()
get_pop_parms()

setCorrelationBlocks(id = list(c("dS", "lBt0", "lp")))

runScenario()
get_pop_parms()

setPopulationParameterInformation(
  a = list(initialValue = getEstimatedPopulationParameters()[["a"]], method = "FIXED")
  )
runScenario()
pre_linearization = get_pop_parms()

scenario <- getScenario()
scenario$linearization = TRUE
setScenario(scenario)
runScenario()
post_lin = get_pop_parms()

full_join(pre_linearization, post_lin, by = "parameter") %>%
  select(parameter, sort(current_vars()))

save_project(vl_dir_tag("holte_vl_refitomegacorr_placebo"))

```

```{r rsif-refit-omegacorr--trt, eval = F}

placebo_refitomegacorr = get_pop_ests(vl_dir_tag("holte_vl_refitomegacorr_placebo")) %>%
  select(parameter, value) %>% spread(parameter, value)
  
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl-vrc01.csv",
  variable_types = vl_types,
  initial_estimates = TRUE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = placebo_refitomegacorr)
setPopulationParameterInformation(
  a = list(initialValue = placebo_refitomegacorr$a, method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )
setCorrelationBlocks(id = list(c("dS", "lBt0", "lp")))
setPopulationParameterInformation(
    corr_lBt0_dS = list(initialValue = placebo_refitomegacorr$corr_lBt0_dS, method = "FIXED"),
    corr_lp_dS = list(initialValue = placebo_refitomegacorr$corr_lp_dS,  method = "FIXED"),
    corr_lp_lBt0 = list(initialValue = placebo_refitomegacorr$corr_lp_lBt0,  method = "FIXED")
  )

stopifnot(nrow(subset(left_join(getPopulationParameterInformation(), 
                                  gather(placebo_refitomegacorr), by = c("name" = "key")), 
                        abs(initialValue - value) > 1e-4 &
                          name != "a" &
                          !is.na(value) )
                 ) == 0
  )
  


runScenario()
save_project(vl_dir_tag("holte_vl_refitomegacorr_vrc01"))

```

## Rerun Dan

```{r dan-lognormal, eval = F}
amp_recipe(
  model_text_file = "Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = vl_types,
  initial_estimates = FALSE,
  obs_dist = "lognormal",
  error_model = "constant"  
  )

set_holte_parms(holte_parms = dan_pop_fits, est_corr_rsif = "FIXED")
setPopulationParameterInformation(
  a = list(initialValue = log(10^dan_pop_fits$a), method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )
getPopulationParameterInformation()

# get initial estimates w/ simulated annealing on
runScenario()

save_project(vl_dir_tag("holte_vl_dan_lognormal"))

```


```{r dan-reproduce, eval = F}
log10vl_types_dan = log10vl_types
log10vl_types_dan[log10vl_types_dan$var_name == "limit", ]$type = "ignore"

amp_recipe(
  model_text_file = "2_Holte_Model.txt",
  data_file = "amp-mlx-vl.csv",
  variable_types = log10vl_types_dan,
  initial_estimates = FALSE,
  obs_dist = "normal",
  error_model = "constant"  
  )
setObservationDistribution(DV_log = "normal")
setErrorModel(DV_log = "constant")

set_holte_parms(holte_parms = dan_pop_fits, est_corr_rsif = "FIXED")
setPopulationParameterInformation(
  a = list(initialValue = dan_pop_fits$a, method = "FIXED"),
  initT_pop = list(method = "FIXED"),
  omega_initT = list(method = "FIXED")
  )
getPopulationParameterInformation()

runScenario()

save_project(vl_dir_tag("holte_vl_dan_rerun"))

```

