---
title: "Infection time assessment"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

Here, we assess the RV217 Holte models to the AMP participants fit using various approaches in fit-vl-model-holte. 

```{r load-packages, echo = T, message=F,output=F, warning=F}
knitr::opts_chunk$set(echo = F)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(GGally)
source(here("R", "directory-funs.R"))
source(here("R", "mlx-model-funs.R"))
source(here("R", "mlx-res-processing.R"))
source(here("R", "mlx-plot-funs.R"))
vl_dir_tag = function(x) paste0("VL/", x)
vl_model_here = function(x, ...) mlx_model_here(vl_dir_tag(x), ...)

theme_set(theme_bw() + theme(legend.position = "top"))
options(scipen = 999)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)

demo_merge = read_csv(clean_data_here("adata-vl.csv"), col_types = cols()) %>%
  dplyr::select(pub_id, contains("gmt"), nisolates, protocol, study, rx_code2) %>% 
  distinct() %>%
  left_join(rx_lab, by = "rx_code2")
stopifnot(nrow(demo_merge) == 158)

```

# Model evaluation

```{r fit-model-setup}

placebo_models = tribble(
  ~model_label, ~model_location,
  "RSIF-placebo", "holte_vl_corr_placebo",
  #"RSIF_nocorr-placebo", "holte_vl_placebo",
  "holte_refit_omega-placebo", "holte_vl_omega_placebo",
  "holte_refit_corr-placebo", "holte_vl_refitcorr_placebo",
  "holte_refit_omega_corr-placebo", "holte_vl_refitomegacorr_placebo", 
)

vrc01_models = tribble(
  ~model_label, ~model_location,
  "RSIF-vrc01", "holte_vl_corr_vrc01",
  #"RSIF_nocorr-vrc01", "holte_vl_vrc01",
  "holte_refit_omega-vrc01", "holte_vl_omega_vrc01",
  "holte_refit_corr-vrc01", "holte_vl_refitcorr_vrc01",
  "holte_refit_omega_corr-vrc01", "holte_vl_refitomegacorr_vrc01", 
)

holte_models = tribble(
  ~model_label, ~model_location,
  "RSIF", "holte_vl_corr_all",
  #"RSIF_nocorr", "holte_vl_all",
  "holte_refit_omega", "holte_vl_omega_all",
  "holte_refit_corr", "holte_vl_refitcorr_all",
  "holte_refit_omega_corr", "holte_vl_refitomegacorr_all"
)

parameter_labels = tribble(
  ~parm, ~parameter_label,
  "aS", "alphaS",
  "log10_dS", "log10(deltaS)",
  "dS", "deltaS",
  "lBt0", "log10(beta)",
  "dI", "kappa",
  "log10_dI", "log10(kappa)",
  "n", "h",
  "lp", "log10(pi)",
  "initT", "First pos.- Inf. day"
)

dan_holte_est = read_csv(here("output/vl-analysis/estimatedIndividualParameters.txt"), col_types = cols()) %>%
  mutate(model = "Dan-Holte") %>%
  select(pub_id = id, model, contains("_mode"), -V0_mode, -aS_mode) %>%
  left_join(demo_merge, by = "pub_id")

paul_times = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols()) %>%
  filter(vl_model_flag) %>%
  mutate(model = "Poisson-fitter", tinf_est = fp_day - tpf_day) %>%
  select(pub_id, model, tinf_est, rx_code2)

pop_ests = stack_model_ests(vl_dir_tag(holte_models$model_location),  
                            holte_models$model_label) 
 
indiv_vl_parms_placebo = stack_indiv_parms(vl_dir_tag(placebo_models$model_location),
                                           placebo_models$model_label) %>%
  select(pub_id = id, model, study_region_cat, contains("_mode"), -V0_mode, -aS_mode) %>%
  left_join(demo_merge, by = "pub_id") 

indiv_vl_parms_vrc01 = stack_indiv_parms(vl_dir_tag(vrc01_models$model_location), 
                                         vrc01_models$model_label) %>%
  select(pub_id = id, model, study_region_cat, contains("_mode"), -V0_mode, -aS_mode) %>%
  left_join(demo_merge, by = "pub_id") 

indiv_vl_parms_all = stack_indiv_parms(vl_dir_tag(holte_models$model_location), 
                                       holte_models$model_label) %>%
  select(pub_id = id, model, study_region_cat, contains("_mode"), -V0_mode, -aS_mode) %>%
  left_join(demo_merge, by = "pub_id") 

```

## Population Parameters

```{r pop-ests}

pop_ests %>%
  dplyr::mutate(parameter = if_else(parameter == "OFV", "LogLik", parameter)) %>%
  dplyr::filter(parameter != "V0_pop") %>%
  left_join(mutate(parameter_labels, parm = paste0(parm, "_pop")), by = c("parameter" = "parm")) %>%
  mutate(
    parameter = if_else(parameter == "OFV", "LogLik", parameter),
    parameter_label = if_else(is.na(parameter_label), "", parameter_label),
    value = if_else(parameter == "dS_pop", value*10^4, value),
    parameter = if_else(parameter == "dS_pop", "dS_pop x 10^-4", parameter),
    estimate_rse = if_else(!is.na(rse_sa),
                           stat_paste(value, abs(rse_sa), digits = 2),
                           stat_paste(value, digits = 2)
  )) %>%
  dplyr::select(model, parameter_label, parameter, estimate_rse) %>% 
  pivot_wider(values_from = estimate_rse, names_from = model)  %>%
  #dplyr::select(parameter_label, parameter, contains("omega"), contains("refit")) %>%
  arrange(desc(parameter_label)) %>%
  kable(format = "html", booktabs = TRUE, linesep = "", escape = FALSE,
        caption = "Population parameters (%RSE). RSEs are not estimated for fixed parameters or the error term.") %>%
  kable_styling(full_width = F)

```

## Evaluate infection times

1. fitted to placebo 
2. two stage for vrc01
3. fitted to all

Consistency between models among 1., 2 and 3
Consistence between models/fits between 1 and 2; and 2 and 3 among placebo or treated

```{r}

indiv_vl_parms_placebo %>%
  bind_rows(indiv_vl_parms_vrc01) %>%
  mutate(model_data = "placebo_trained", model = str_remove(str_remove(model, "-placebo"), "-vrc01")) %>%
  bind_rows(mutate(subset(indiv_vl_parms_all), model_data = "full_model")) %>%
  select(pub_id, tinf_est = initT_mode, model, model_data) %>%
  ggplot(aes(x = model_data, y = tinf_est)) +
  geom_boxplot() +
  ylab("est. inf day (prior to first positive)") +
  geom_path(aes(group = pub_id), alpha = 0.15) +
  facet_wrap( ~ model, nrow = 1)


```

### Consistency within trt groups

```{r}

indiv_vl_parms_placebo %>%
  mutate(model_data = "placebo_only", model = str_remove(model, "-placebo")) %>%
  bind_rows(mutate(subset(indiv_vl_parms_all, rx_code2 == "C"), model_data = "full_model")) %>%
  select(pub_id, tinf_est = initT_mode, model, model_data) %>%
  ggplot(aes(x = model_data, y = tinf_est)) +
  geom_boxplot() +
  ylab("est. inf day (prior to first positive)") +
  geom_path(aes(group = pub_id), alpha = 0.15) +
  facet_wrap( ~ model)

indiv_vl_parms_placebo %>% 
  bind_rows(subset(indiv_vl_parms_all, rx_code2 == "C")) %>%
  select(pub_id, tinf_est = initT_mode, model) %>%
  pivot_wider(values_from = tinf_est, names_from = model) %>%
  dplyr::select(-pub_id) %>% 
  ggpairs(progress = F, xlab = "First pos - inf time",
          ylab = "First pos - inf time")

```

```{r}

indiv_vl_parms_vrc01 %>%
  mutate(model_data = "placebo_trained", model = str_remove(model, "-vrc01")) %>%
  bind_rows(mutate(subset(indiv_vl_parms_all, rx_code2 != "C"), model_data = "full_trained")) %>%
  select(pub_id, tinf_est = initT_mode, model, model_data) %>%
  ggplot(aes(x = model_data, y = tinf_est)) +
  geom_boxplot() +
  ylab("est. inf day (prior to first positive)") +
  geom_path(aes(group = pub_id), alpha = 0.15) +
  facet_wrap( ~ model)

indiv_vl_parms_vrc01 %>% 
  bind_rows(subset(indiv_vl_parms_all, rx_code2 != "C")) %>%
  select(pub_id, tinf_est = initT_mode, model) %>%
  pivot_wider(values_from = tinf_est, names_from = model) %>%
  dplyr::select(-pub_id) %>% 
  ggpairs(progress = F, xlab = "First pos - inf time",
          ylab = "First pos - inf time")

```


```{r}


indiv_vl_parms_placebo %>%
  bind_rows(indiv_vl_parms_vrc01) %>%
  mutate(model = str_remove(str_remove(model, "-placebo"), "-vrc01")) %>%
  bind_rows(dan_holte_est) %>%
  dplyr::select(pub_id, tinf_est = initT_mode, model) %>%
  bind_rows(select(paul_times, -rx_code2)) %>%
  pivot_wider(values_from = tinf_est, names_from = model) %>%
  dplyr::select(-pub_id) %>% 
  ggpairs(progress = F, xlab = "First pos - inf time",
          ylab = "First pos - inf time")

```

## Generate fits

```{r inf-time-dist-placebo}

indiv_vl_parms_placebo %>%
  bind_rows(subset(dan_holte_est, rx_code2 == "C")) %>%
  select(pub_id, tinf_est = initT_mode, model) %>%
  bind_rows( bind_rows(subset(paul_times, rx_code2 == "C"))) %>%
  ggplot(aes(x = model, y = tinf_est)) +
  geom_boxplot() +
  ylab("est. inf day (prior to first positive)") +
  geom_path(aes(group = pub_id), alpha = 0.15)

```

```{r inf-time-dist}

indiv_vl_parms %>%
  bind_rows(dan_holte_est) %>%
  select(pub_id, tinf_est = initT_mode, model) %>%
  bind_rows(paul_times) %>%
  ggplot(aes(x = model, y = tinf_est)) +
  geom_boxplot() +
  ylab("est. inf day (prior to first positive)") +
  geom_path(aes(group = pub_id), alpha = 0.15)

```

```{r tinf-correlations-placebo}

```


```{r tinf-correlations}

indiv_vl_parms %>% 
  bind_rows(dan_holte_est) %>%
  dplyr::select(pub_id, tinf_est = initT_mode, model) %>%
  bind_rows(paul_times) %>%
  pivot_wider(values_from = tinf_est, names_from = model) %>%
  dplyr::select(-pub_id) %>% 
  ggpairs(progress = F, xlab = "First pos - inf time",
          ylab = "First pos - inf time")

```


```{r pop-fits}

stack_model_ests(time_models) %>%
  dplyr::mutate(parameter = if_else(parameter == "OFV", "LogLik", parameter)) %>%
  dplyr::filter(parameter != "V0_pop") %>%
  left_join(mutate(parameter_labels, parm = paste0(parm, "_pop")), by = c("parameter" = "parm")) %>%
  left_join(time_model_names, by = "model") %>%
  mutate(
    parameter = if_else(parameter == "OFV", "LogLik", parameter),
    parameter_label = if_else(is.na(parameter_label), "", parameter_label),
    value = if_else(parameter == "dS_pop", value*10^4, value),
    parameter = if_else(parameter == "dS_pop", "dS_pop x 10^-4", parameter),
    estimate_rse = if_else(!is.na(rse_sa),
                           stat_paste(value, abs(rse_sa), digits = 2),
                           stat_paste(value, digits = 2)
  )) %>%
  dplyr::select(model_label, parameter_label, parameter, estimate_rse) %>% 
  pivot_wider(values_from = estimate_rse, names_from = model_label)  %>%
  #dplyr::select(parameter_label, parameter, contains("omega"), contains("refit")) %>%
  arrange(desc(parameter_label)) %>%
  kable(format = "html", booktabs = TRUE, linesep = "", escape = FALSE,
        caption = "Population parameters (%RSE). RSEs are not estimated for fixed parameters or the error term.") %>%
  kable_styling(full_width = F)

```

## Correlation plots

#### Raw plot with spearman cor.

```{r time-est, fig.height=9, fig.width=9, warning = F, message = F}
time_ests %>%
  dplyr::filter(placebo) %>%
  dplyr::select(-model, -placebo) %>%
  pivot_wider(values_from = tinf_est, names_from = model_label) %>%
  dplyr::select(-id) %>% 
  ggpairs(progress = F, xlab = "First pos - inf time",
           ylab = "First pos - inf time") 

```


#### Same plot -  Lin's CCC displayed. Axes have common limits with 1:1 line (no TCL free, all = 15 days) .

```{r time-est-scaled, fig.height=9, fig.width=9, warning = F, message=F}
my_custom_cor <- function(data, mapping, sizeRange = c(1, 5), ...) {

  # get the x and y data to use the other code
 
  x <-  GGally::eval_data_col(data, mapping$x)
  y <-  GGally::eval_data_col(data, mapping$y)

  if(sd(x) == 0 | sd(y) == 0){
    tt = "NA"
  } else{
    ct <- epiR::epi.ccc(x, y)
    r <- unname(ct$rho.c$est)
    rt <- format(r, digits = 2)[1]
    tt <- as.character(rt)
  }
  #tt = "test"

  # plot the cor value
  p <- ggally_text(
   label = tt, 
   mapping = aes(),
   xP = 40, yP = 40, 
   size = 4,
   ...
  ) +

theme(panel.background=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major=element_blank()) 

 p
}

my_custom_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point() + 
    scale_color_brewer(palette="Accent") +
    geom_abline(color = I("black")) +
    theme(panel.background=element_blank(),
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(color="gray"))
}


my_boxplot = function (data, mapping, ..., boxPlot = TRUE) {
    ggplot(data = data, aes(y = 40))  + geom_boxplot(mapping, ..., width = 30) +
    coord_flip()
}

suppressMessages({
time_ests %>%
  dplyr::filter(placebo & model_label != "TCL_free") %>%
  dplyr::select(-model, -placebo) %>%
  pivot_wider(values_from = tinf_est, names_from = model_label) %>%
  dplyr::select(-id) %>% 
  ggpairs(upper = list(continuous = wrap(my_custom_cor)), 
          lower = list(continuous = my_custom_smooth),
          diag = list(continuous = my_boxplot),
          progress = F, xlab = "First pos - inf time",
           ylab = "First pos - inf time") +
  scale_x_continuous(limits = c(2.5, 75)) +
  scale_y_continuous(limits = c(2.5, 75))
})

```

## Individual trajectories

**colored by model, dots are the estimated infection time (t_det). Open black dot is Paul's estimate. Raw data subsetted to at least 100 weeks within first positive**

```{r indiv-traj-setup}

fits = stack_indiv_fits(time_models) %>%
  mutate(weeks = time / 7) %>%
  left_join(time_ests, by = c("model", "ID" = "id")) %>%
  dplyr::filter(ID %in% placebo_ids & time >= -tinf_est)

ipreds = stack_indiv_ests(time_models) %>%
  mutate(weeks = time/7) %>%
  dplyr::filter(ID %in% placebo_ids)

```


```{r traj-test-pl, include = F, eval = F}
test_set = placebo_ids[1:8]

test_fits = stack_indiv_fits(time_models) %>%
  mutate(weeks = time / 7) %>%
  left_join(time_ests, by = c("model", "ID" = "id")) %>%
  dplyr::filter(ID %in% test_set & time >= -tinf_est)

test_ipreds = stack_indiv_ests(time_models) %>%
  mutate(weeks = time/7) %>%
  dplyr::filter(ID %in% test_set)


ggplot(test_fits, aes(x = time, y = log10(pmax(0.01, indivPredMode)))) +
  geom_line(aes(colour = model_label)) +
  geom_point(data = test_ipreds, aes(y = log10(DV), shape = as.factor(censored != 0))) +
  geom_point(data = rename(dplyr::filter(time_ests, id %in% test_set & !is.na(model)), ID = id),
             aes(x = -tinf_est, y = log10(0.01), colour = model_label)) +
  geom_point(data = rename(dplyr::filter(paul_times, id %in% test_set), ID = id),
             aes(x = -tinf_est, y = log10(0.01)), shape = 1) +  
  scale_shape_discrete("Below LLoQ") +
  facet_wrap(~ ID, scales = "free_x") +
  labs(y = "log10 VL", colour = "") +
  theme(legend.position = 'top')+ 
  guides(shape = guide_legend(ncol = 1, nrow = 2))

```


```{r indiv-pl}
#| message = FALSE, warning=F,
#| fig.width=10, fig.height=10,

ggplot(fits, aes(x = time, y = log10(pmax(0.01, indivPredMode)))) +
  geom_line(aes(colour = model_label)) +
  geom_point(data = dplyr::filter(ipreds, time > -100),
             aes(y = log10(DV), shape = as.factor(censored != 0))) +
  geom_point(data = rename(dplyr::filter(time_ests, placebo & !is.na(model)), ID = id),
             aes(x = -tinf_est, y = log10(0.01), colour = model_label)) +
  geom_point(data = rename(dplyr::filter(paul_times, id %in% placebo_ids), ID = id),
             aes(x = -tinf_est, y = log10(0.01)), shape = 1) +  
  scale_shape_discrete("Below LLoQ") +
  facet_wrap(~ ID, nrow = 9, ncol = 7, scales = "free_x") +
  labs(y = "log10 VL", colour = "", x = "weeks") +
  theme(legend.position = 'top', axis.text.x = element_text(size = 8)) + 
  guides(shape = guide_legend(ncol = 1, nrow = 2))

```
