---
title: "Results from final AMP VL models"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

This curates the final results from the model building:

- The AIC table showing best models for placebo and VRC01 groups
  - Selection of "final" models for placebo and VRC01 cohorts
  - the final adjusted results for placebo just has indirect effects
  - the final adjusted results for vrc01 has both
- Look at final model vs. unadjusted results
  - parameter comparisons
  - trajectories
  - vl summary comparisons

Using the final models, simulate the VL. save the key parameters, R0, calculate key model sim results (AUC, set point, upslope). Merge on the observed vl summary.

The following summaries are calculated from model simulations, and claculated by `summarize_sim_holte_vl()` from `summarize_sim_holte_vl.R`

- peak: maximum log10vl
- peak day
- 3-mo setpoint: avg log10vl betweens days 90 and 100
- final setpoint: avg log10vl between days 950 and 1000 (not for analysis)
- 3-month AUC/t: auc of vl through day 90 (linear up, log down method), divided by 90
- 3-month geometric AUC: auc of log10 vl through day 90 divided by 90 (trapezoid)
- upslope: max slope of vl rise between days 0 and (5, 21)
- upslope R0: (1+upslope/dI):  (Ribeiro, PMC2876646)
- 'unadjusted' R0: beta * pi * aS / (dS * dI * 23)

This was already done for the unadjusted model in 02_analysis-vl-model-holte.Rmd: `unadjusted_model_vl_summary.csv`

```{r load-packages, echo = T, message=F,output=F, warning=F}
knitr::opts_chunk$set(echo = T)
library(VISCfunctions)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(GGally)
# library(pkr) use but dont load
source(here("R", "directory-funs.R"))
source(here("R", "mlx-res-processing.R"))
source(here("R", "summarize_sim_holte_vl.R"))
source(here("R", "rxode-funs.R"))
source(here("models", "rxode-models", "rxode-models.R"))
source(here("models", "rxode-models", "rxode-pkpd-models.R"))

hill_slope = function(ic50, ic80) -log(4)/(log(ic50/ic80))

knitr::opts_chunk$set(dev = c("png", "pdf"))

strip_as_yaxis = theme(strip.placement = "outside", strip.background.y = element_blank())
theme_set(theme_bw() + theme(legend.position = "top"))
options(scipen = 999)

rx_lab = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat = c("Placebo", "VRC01 (10mg/kg)", "VRC01 (30mg/kg)")
)
rx_lab_pool = tibble(
  rx_code2 = c("C", "T1", "T2"),
  trt_cat_pool = c("Placebo", "VRC01 (Pooled)", "VRC01 (Pooled)")
)

ic80gte1_colors = c(`IC80 < 1 mcg/mL` = "#008080", `IC80 >= 1 mcg/mL` = "coral")

save_shared_folder = T

```


```{r load-raw-data}
## placebo is the indirect model
## vrc01 is the full indirect+pkpd model
final_placebo_loc = "VL-adjusted-placebo/cns-pi"
final_vrc01_loc = "PKPD/adj-pkpd-model-no-omega"

time_data = read_csv(clean_data_here("adata-time-summary.csv"), col_types = cols()) %>%
  select(pub_id, contains("_day"))

demo_merge_full = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols()) %>%
  dplyr::select(pub_id, contains("gmt"), nisolates, protocol, study, rx_code2, vl_model_flag) %>% 
  distinct() %>%
  left_join(rx_lab, by = "rx_code2") %>%
  mutate(
    hill_ls = hill_slope(gmt50ls, gmt80ls),
    hill_ms = hill_slope(gmt50ms, gmt80ms),
    hill_mf = hill_slope(gmt50mf, gmt80mf),
    hill_ls = if_else(gmt80ls < 100, hill_ls, mean(hill_ls[gmt80ls < 100 & rx_code2 != "C"])),
    hill_ms = if_else(gmt80ms < 100, hill_ms, mean(hill_ms[gmt80ms < 100 & rx_code2 != "C"])),
    hill_mf = if_else(gmt80mf < 100, hill_mf, mean(hill_mf[gmt80mf < 100 & rx_code2 != "C"])),
  ) %>%
  left_join(time_data, by = "pub_id") 

demo_merge = filter(demo_merge_full, vl_model_flag) %>% select(-vl_model_flag)

stopifnot(nrow(demo_merge) == 158)

# raw data
all_indiv_obs = get_indiv_preds("VL/holte_vl_corr_all") %>%
  rename(pub_id = ID) %>%
  left_join(select(time_data, pub_id, fp_day), by = "pub_id") %>%
  mutate(
    time = time + fp_day,
    weeks = time/7, 
    log10V = log10(DV)) %>%
  select(pub_id, time, weeks, fp_day, DV, log10V, censored)

vrc01_initT = read_csv(clean_data_here("adjusted_vrc01_parms.csv"), col_types = cols()) %>%
  select(pub_id, initT_mode)

# pk data

pk_pop_ests = get_pop_ests("PK/VRC01AMP_Case-Cntrl") %>%
  dplyr::select(-se_sa, -rse_sa) %>%
  spread(key = parameter, value = value)

vrc01_pkpd_dat = read_csv(mlx_data_here("amp-mlx-pkpd-indirect-holteT0.csv"), col_types = cols())

vrc01_pk_parms = distinct(vrc01_pkpd_dat, pub_id, Cl, V1, Q, V2)

mlx_infection_times = vrc01_pkpd_dat %>% filter(ADM == 2) %>% select(pub_id, infection_day = TIME)

vrc01_dose = vrc01_pkpd_dat %>%
  filter(ADM == 1) %>%
  distinct(pub_id, TIME, AMT, RATE) %>%
  mutate(AMT = as.numeric(AMT), RATE = as.numeric(RATE)) %>%
  arrange(pub_id, TIME) %>%
  group_by(pub_id) %>%
  mutate(DOSENO = row_number()) %>%
  ungroup()

```

```{r summarize-obs-data}
vl_data = read_csv(clean_data_here("adata-vl-stats.csv"), col_types = cols()) 

vl_data_summary = vl_data %>%
  group_by(pub_id, vl_model_flag) %>%
  summarize(
    first_pos = log10vl[days_fp == 0],
    n_avg = sum(days_fp >= 0),
    avg_vl = mean(log10vl[days_fp >= 0]),
    peak_vl = max(log10vl),
    peak_day_fp = days_fp[which.max(log10vl)],
    n_setpt = sum(days_fp > days_fp[which.max(log10vl)]),
    setpt = if_else(n_setpt == 0, NA_real_, mean(log10vl[days_fp > days_fp[which.max(log10vl)]])),
    .groups = "drop"
  )
stopifnot(nrow(vl_data_summary) == 162)

```

```{r load-pop-parms-fits}

# builds upon placebo model
final_pop_parms = get_model_ests(final_vrc01_loc) 
final_pop_theta = prep_vl_model_parms(final_vrc01_loc)

final_mlx_fits = stack_indiv_fits(c(final_placebo_loc, final_vrc01_loc)) %>%
  rename(pub_id = ID) %>%
  left_join(demo_merge, by = "pub_id") %>%
  mutate(
    time = if_else(rx_code2 == "C", time + fp_day, time), #placebo model scaled to first positive
    weeks = time / 7,
    log10V = log10(pmax(20, indivPredMode)),
    model = "mlx-predict"
    ) %>%
  left_join(mlx_infection_times, by = "pub_id")

```

```{r setup-indiv-parms}

# from unadjusted model
unadj_vl_summary = read_csv(clean_data_here("unadjusted_model_vl_summary.csv"), col_types = cols()) %>%
  mutate(
    infection_time = fp_day - initT_mode,
    model = 'unadjusted')

# rest combining final models

# just add pop PK parms
final_placebo_parms = get_indiv_parms(final_placebo_loc) %>%
  rename(pub_id = id) %>%
  mutate(
    study = if_else(str_detect(study_region_cat, "704"), 1, 0),
    Cl = exp(log(pk_pop_ests$Cl_pop) + pk_pop_ests$beta_Cl_study_1*(study == 1)),
    V1 = pk_pop_ests$V1_pop,
    Q = pk_pop_ests$Q_pop,
    V2 = exp(log(pk_pop_ests$V2_pop) + pk_pop_ests$beta_V2_study_1 * (study == 1))
  ) %>%
  select(-study)

final_vrc01_parms = get_indiv_parms(final_vrc01_loc) %>%
  rename(pub_id = id) %>%
  left_join(vrc01_initT, by = "pub_id") %>%
  left_join(vrc01_pk_parms, by = "pub_id")

# explicitly lrho_mode code with "C" to catch other errors
final_vl_parms = bind_rows(final_placebo_parms, final_vrc01_parms) %>%
  select(pub_id, contains("_mode"), study_region_cat, Cl, V1, Q, V2) %>%
  left_join(demo_merge, by = "pub_id") %>%
  mutate(
    infection_time = if_else(rx_code2 != "C", floor(fp_day - initT_mode), fp_day - initT_mode), # former matches data setup
    lrho_mode = if_else(rx_code2 == "C" & is.na(lrho_mode), final_pop_theta$lrho, lrho_mode),
    V0_mode = 0.01
  )

stopifnot(nrow(final_vl_parms) == 158)
stopifnot(all(!is.na(select(final_vl_parms))))

```

# Population Parameters

```{r pop-parms-setup}
all_models = tribble(
  ~model_location, ~model_label, ~paper_model,
  "VL/holte_vl_corr_placebo", "Placebo_Fixed RV217_(Reeves et al., RSIF 2021)", T,
  "VL/holte_vl_refitomega_placebo", "Placebo_Unadjusted_Refit RV217 Variance", F,
  "VL/holte_vl_refitcorr_placebo", "Placebo_Unadjusted_Refit RV217 Correlation", T,
  "VL/holte_vl_refitomegacorr_placebo", "Placebo_Unadjusted_Refit RV217 Variance/Correlation", F,
  "VL-adjusted-placebo/adj-beta", "Placebo_Adjusted_Cat. IC80-adjusted infectivity", F,
  "VL-adjusted-placebo/adj-pi", "Placebo_Adjusted_Cat. IC80-adjusted viral burst", F,
  "VL-adjusted-placebo/adj-kappa", "Placebo_Adjusted_Cat. IC80-adjusted infected cell death", F,
  "VL-adjusted-placebo/adj-n", "Placebo_Adjusted_Cat. IC80-adjusted Holte power", F,
  "VL-adjusted-placebo/cns-beta", "Placebo_Adjusted_IC80-adjusted infectivity", F,
  "VL-adjusted-placebo/cns-pi", "Placebo_Adjusted_IC80-adjusted viral burst", T,
  "VL-adjusted-placebo/cns-kappa", "Placebo_Adjusted_IC80-adjusted infected cell death", F,
  "VL-adjusted-placebo/cns-n", "Placebo_Adjusted_IC80-adjusted Holte power", F,
  "VL/holte_vl_refitcorr_vrc01", "VRC01_PKPD_Unadjusted (RV217 refit correlation)", F,
  "VL-adjusted-placebo/cns-pi-vrc01", "VRC01_PKPD_Indirect only (IC80-adjusted viral burst)", F,
  "PKPD/pkpd-model", "VRC01_PKPD_PKPD VRC01-adjusted only (w/ potency reduction variance)", F,
  "PKPD/pkpd-model-no-omega", "VRC01_PKPD_PKPD VRC01-adjusted only", F,
  "PKPD/adj-pkpd-model", "VRC01_PKPD_Indirect + PKPD VRC01-adjusted (w/ potency reduction variance)", F,
  "PKPD/adj-pkpd-model-no-omega", "VRC01_PKPD_Indirect + PKPD VRC01-adjusted", T
  )

all_pop_est = stack_model_ests(all_models$model_location, all_models$model_label) %>%
  left_join(all_models, by = c("model" = "model_label")) %>%
  separate(model, into = c("Cohort", "Virus-adjusted", "Model Description"), sep = "_", remove = F) %>%
  mutate(
    parameter = if_else(parameter == "OFV", "LogLik", parameter),
    `Virus-adjusted` = factor(`Virus-adjusted`,
                              levels = c("Fixed RV217", "Unadjusted", "Adjusted", "PKPD"))
  ) 
```

## Model AIC Comparisons

```{r bic-table}

all_pop_est %>%
  filter(parameter == "LogLik" | parameter == "BICc") %>% 
  select(-se_sa, -rse_sa, -model) %>%
  spread(parameter, value) %>%
  rename(`Analysis Model` = paper_model) %>%
  arrange(`Virus-adjusted`, -BICc) %>%
  write_csv(here("output/pkpd-modeling/model-information-criteria.csv")) %>%
  kable(format = "html", booktabs = TRUE, linesep = "", escape = FALSE,
        caption = "Model information criteria") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(1:2, valign = "top")

```

## Model building population parameters

```{r pop-parms}

all_pop_est %>%
  mutate(
    keep =  model == "Placebo_Fixed RV217_(Reeves et al., RSIF 2021)" | !is.na(rse_sa)
  ) %>%
  filter(paper_model & keep & !parameter %in% c("LogLik","OFV", "BIC", "BICc", "AIC", "V0_pop")) %>%
  select(-`Virus-adjusted`, -model_location, -model) %>%
  dplyr::mutate(
    parameter = fct_relevel(factor(parameter, levels = unique(all_pop_est$parameter)), "lrho_pop", after = 7),
    parameter = fct_relevel(parameter, "beta_lp_log_ic80_ctr", after = 5),
    value =  if_else(parameter == "dS_pop", value * 10000, value),
    estimate_rse = if_else(!is.na(rse_sa) & `Model Description` != "(Reeves et al., RSIF 2021)",
                           stat_paste(value, abs(rse_sa), digits = 2),
                           stat_paste(value, digits = 2)) 
    ) %>%
  arrange(parameter) %>%
  mutate(
    parameter = str_remove(as.character(parameter), "_pop"),
    parameter =  if_else(parameter == "dS", "dS * 10^-4", parameter)
  ) %>%
  dplyr::select(`Model Description`, parameter, estimate_rse) %>% 
  pivot_wider(values_from = estimate_rse, names_from = `Model Description`)  %>%
  write_csv(here("output/pkpd-modeling/pop-model-parms.csv")) %>%
  kable(format = "html", booktabs = TRUE, linesep = "", escape = FALSE,
        caption = "Fitted population model parameters (%RSE) at each model building step. Empty cells indicate parameter value taken from previous model step.") %>%
  kable_styling(full_width = F)

```

# Simulate VL

```{r rxode-sim}

unadj_vl_sims = map_df(1:nrow(unadj_vl_summary), function(i){
  parm_set = unadj_vl_summary[i, ]
  stopifnot(nrow(parm_set) == 1)
  run_vl_model(c(seq(0, 14, by = 0.1), 15:225), holte_model, 
               theta = prep_vl_model_parms(mode_parms = parm_set)) %>%
  mutate(
    pub_id = parm_set$pub_id, 
    time = time + parm_set$fp_day - parm_set$initT_mode,
    weeks = time/7,
    model = "unadjusted", 
    fp_day = parm_set$fp_day,
    infection_day = parm_set$fp_day - parm_set$initT_mode
    )
})

vl_sims = map_df(1:nrow(final_vl_parms), function(i){
  parm_set = final_vl_parms[i, ]
  stopifnot(nrow(parm_set) == 1)
  
  dosing = filter(vrc01_dose, pub_id == parm_set$pub_id)
  if(nrow(dosing) == 0) dosing = NULL
  
  infection_day = parm_set$infection_time
  mtime = c(0:(infection_day - 1), 
            seq(infection_day, infection_day + 50, by = 0.1), 
            infection_day + 51:225)
  theta_pk = select(parm_set, Cl, V1, Q, V2)
  theta_pd = tibble(IC80 = parm_set$gmt50ls * 4 ^(1/parm_set$hill_ls), 
                    rho = 10^parm_set$lrho_mode, 
                    h = parm_set$hill_ls)
  
  run_pkpd_models(mtime,      
                  holte_pkpd_model,
                  theta_pk = theta_pk,
                  theta_vl = prep_vl_model_parms(mode_parms = parm_set),
                  theta_pd = theta_pd,
                  infection_time = infection_day,
                  infusion_pk_dosing = dosing) %>%
  mutate(
    fp_day = parm_set$fp_day,
    infection_day = infection_day,
    pub_id = parm_set$pub_id, 
    weeks = time/7, 
    model = "final")
})

all_sims = bind_rows(unadj_vl_sims, vl_sims, final_mlx_fits)

```

```{r save-fits}

placebo_ids = subset(final_vl_parms, rx_code2 == "C")$pub_id

# these all look great
indiv_plots_placebo = map(split(sort(placebo_ids),
                         ceiling(seq_along(placebo_ids)/(5 * 2))), 
                  function(id_set) {
                    obs_dat = dplyr::filter(all_indiv_obs, pub_id %in% id_set)
                    all_sims %>%
                      dplyr::filter(pub_id %in% id_set) %>%
                      ggplot(aes(x = weeks - fp_day/7, y = log10V)) +
                      geom_line(aes(colour = model)) +
                      geom_point(data = obs_dat, aes(shape = as.factor(censored != 0))) +
                      scale_shape_discrete(guide = "none") +
                      facet_wrap(~ pub_id, nrow = 5, ncol = 2) +
                      coord_cartesian(xlim = c(-5, NA)) +
                      labs(y = "log10 VL") +
                      theme(legend.position = 'top')
                  })

ggsave(here("output/pkpd-modeling/final-placebo-fits.pdf"), 
       width = 8, height = 10,
       gridExtra::marrangeGrob(indiv_plots_placebo, nrow=1, ncol=1))


vrc01_ids = subset(final_vl_parms, rx_code2 != "C")$pub_id
indiv_plots_vrc01 = map(split(sort(vrc01_ids),
                         ceiling(seq_along(vrc01_ids)/(5 * 2))), 
                  function(id_set) {
                    obs_dat = dplyr::filter(all_indiv_obs, pub_id %in% id_set)
                    all_sims %>%
                      dplyr::filter(pub_id %in% id_set) %>%
                      ggplot(aes(x = weeks - fp_day/7, y = log10V)) +
                      geom_line(aes(colour = model)) +
                      geom_point(data = obs_dat, aes(shape = as.factor(censored != 0))) +
                      scale_shape_discrete(guide = "none") +
                      facet_wrap(~ pub_id, nrow = 5, ncol = 2) +
                      coord_cartesian(xlim = c(-5, NA)) +
                      labs(y = "log10 VL") +
                      theme(legend.position = 'top')
                  })

ggsave(here("output/pkpd-modeling/final-vrc01-fits.pdf"), 
       width = 8, height = 10,
       gridExtra::marrangeGrob(indiv_plots_vrc01, nrow=1, ncol=1))

```

# Model Summary Measures

```{r}

# pk is not relevant
final_setpt = map_df(1:nrow(final_vl_parms), function(i){
  parm_set = final_vl_parms[i, ]
  stopifnot(nrow(parm_set) == 1)
  run_vl_model(950:1000, holte_model, 
               theta = prep_vl_model_parms(mode_parms = parm_set)) %>%
    mutate(pub_id = parm_set$pub_id)
}) %>%
  group_by(pub_id) %>%
  summarize(final_setpt = mean(log10V), .groups = "drop")


vl_summary = vl_sims %>%
  filter(time >= infection_day) %>%
  mutate(time = time - infection_day) %>%
  summarize_sim_holte_vl(vl_sims_dat = ., 
                                    vl_parms = rename(final_vl_parms, model_infection_day = infection_time),
                                    setpt_end = 90, auc_end = 90, 
                                    max_uplope_time = c(3, 21)) %>%
  left_join(final_setpt, by = "pub_id")

stopifnot(all(!is.na(vl_summary)))
stopifnot(nrow(vl_summary) == 158)

```

```{r plot-endpoints, fig.height = 8, fig.width = 7.5}

key_measures = c("upslope_r0", "unadjusted_r0", "model_log10peak", "geo_auc_3mo", "log10_auc_3mo", "model_peak_day",
                 "model_setpt")
ic80gte1_colors = c(`IC80 < 1 mcg/mL` = "#008080", `IC80 >= 1 mcg/mL` = "coral")

(vl_summary %>%
  left_join(rx_lab_pool, by = "rx_code2") %>%
  mutate(
    ic80gte1 = factor(gmt80ls >= 1, levels = c(F, T), labels = c("IC80 < 1 mcg/mL", "IC80 >= 1 mcg/mL"))
    ) %>%
  select(pub_id, key_measures, ic80gte1, trt_cat_pool) %>%
  pivot_longer(names_to = "measure", values_to = "value", cols = key_measures) %>%
  ggplot(aes(x = trt_cat_pool, y = value, colour = ic80gte1)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.75)) +
  facet_wrap(~measure, strip.position = "left", scales = "free_y") +
  ylab("") + xlab("") +
  scale_color_manual("", values = ic80gte1_colors) +
  strip_as_yaxis +
    theme(legend.direction = "vertical")) %>%
  lemon::reposition_legend(position =  'center',  panel = 'panel-3-2')

```

## Measure comparisons

```{r setpt-comp}

vl_summary %>%
  dplyr::select(contains("setpt")) %>%
  ggpairs(progress = F)

```

```{r setpt-models, warning=F,message=F}

unadj_vl_summary %>%
  rename(unadj_model_setpt = model_setpt) %>%
  left_join(vl_summary, by = "pub_id") %>%
  left_join(vl_data_summary, by = "pub_id") %>%
  select(unadj_model_setpt, model_setpt, obs_setpt = setpt)%>%
  ggpairs(progress = F)

```


```{r auc-comp}

vl_summary %>%
  dplyr::select(contains("auc")) %>%
  ggpairs(progress = F)

```

```{r avg-models, warning=F,message=F}

unadj_vl_summary %>%
  rename(unadj_geo_auc_3mo = geo_auc_3mo) %>%
  left_join(vl_summary, by = "pub_id") %>%
  left_join(vl_data_summary, by = "pub_id") %>%
  select(unadj_geo_auc_3mo, geo_auc_3mo, obs_avg = avg_vl)%>%
  ggpairs(progress = F)

```

```{r r0-comp}

vl_summary %>%
  dplyr::select(contains("r0")) %>%
  ggpairs(progress = F)

```


```{r r0-models, warning=F,message=F}

unadj_vl_summary %>%
  rename(unadj_upslope_r0 = upslope_r0, unadj_calc_r0 = unadjusted_r0) %>%
  left_join(rename(vl_summary, calc_r0 = unadjusted_r0), by = "pub_id") %>%
  dplyr::select(contains("r0")) %>%
  ggpairs(progress = F)

```

# Save final 


```{r}

non_model_ptids = vl_data_summary %>%
  filter(!vl_model_flag) %>%
  select(-vl_model_flag) %>%
  left_join(demo_merge_full, by = "pub_id")

stopifnot(nrow(non_model_ptids) == 4)

```

```{r, class.source = "fold-show"}

write_csv(final_pop_parms, clean_data_here("final_model_popparms.csv"))

final_vl_summary = left_join(vl_summary, vl_data_summary,  by = "pub_id") %>%
  bind_rows(non_model_ptids)

# 4 ptids really just have first positive data
stopifnot(nrow(final_vl_summary) == 162)
stopifnot(all(!is.na(select(subset(final_vl_summary, !is.na(initT_mode)), -setpt))))
stopifnot(nrow(subset(final_vl_summary, !is.na(initT_mode))) == 158)

glimpse(final_vl_summary)
write_csv(final_vl_summary, clean_data_here("final_vl_summary.csv"))

```

```{r save-shared, eval = save_shared_folder}

write_csv(tibble(), 
           paste0("/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/!FINAL VL MODEL UPDATED - ",
                  Sys.Date()))

write_csv(select(final_vl_summary, -log10_auc_3mo, -study, -final_infusion_pre_fp_day, -final_infusion_pre_dx_day),  
          "/Volumes/trials/vaccine/Bryan_Mayer/AMP-Modeling/data/final_model_vl_summary.csv")

```
