---
title: "Simulating PKPD using AMP settings"
author: "Bryan Mayer"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: true
    code_folding: hide
---

# Overview

This is to generate the final figures from the PKPD model decsribing the population results in
the AMP study.

- Placebo effect plot (the burst size adjustment model)
- PKPD plots showing direct and indirect effects


```{r load-packages, echo = T, message=F,output=F,warning =F}
knitr::opts_chunk$set(echo = T)
library(tidyverse)
library(here)
library(glue)
library(kableExtra)
library(RxODE)
library(viridis)
library(cowplot)
source(here("R", "directory-funs.R"))
source(here("R", "mlx-res-processing.R"))
source(here("R", "rxode-funs.R"))
source(here("models", "rxode-models", "rxode-models.R"))
source(here("models", "rxode-models", "rxode-pkpd-models.R"))

theme_set(theme_bw() + theme(legend.position = "top"))
strip_as_yaxis = theme(strip.placement = "outside", strip.background.y = element_blank())

titer_neut = function(id50, h = 1) 1 - beta_titer_neut(id50, h)
beta_titer_neut = function(id50, h = 1) 1/(1+id50^h)

```

```{r amp-pe-data}

pt80_efficacy = tibble(
  PT80 = c(32, 82, 194),
  PE = c(50, 75, 90),
  PE_pct = paste0("PE ", PE, "%"),
  pe_neut = PE/100
)

empirical_amp_fit = drc::drm(pe_neut ~ PT80, data = pt80_efficacy, 
                        fct = drc::LL.5(names = c("slope", "lower", "upper",
                                                  "inflection", "asymmetry"),
                             fixed = c(NA, 0, 1, NA, NA)))

```


```{r load-placebo-mod}
raw_placebo_data = read_csv(mlx_data_here("amp-mlx-vl-placebo.csv"), col_types = cols())

placebo_ic80_range = c(min(raw_placebo_data$gmt80ls), max(raw_placebo_data$gmt80ls))
placebo_ic50_range = c(min(raw_placebo_data$gmt50ls), max(raw_placebo_data$gmt50ls))

final_pop_ests = read_csv(clean_data_here("final_model_popparms.csv"), col_types = cols())
final_pop_ests_wide = final_pop_ests %>%
  dplyr::select(parameter, value) %>% spread(parameter, value)

theta_vl_placebo = prep_vl_model_parms(mode_parms = final_pop_ests_wide, parm_tag = "_pop")

burst_effect = function(ic80, pop_parms = placebo_pop_wide) {
  lp = final_pop_ests_wide$lp_pop + final_pop_ests_wide$beta_lp_log_ic80_ctr * (log10(ic80) - 0.638)
  10^(lp)
}

```

```{r load-trt-mod}

raw_vrc01_data = read_csv(mlx_data_here("amp-mlx-vl-vrc01.csv"), col_types = cols())

vrc01_ic80_range = c(min(raw_vrc01_data$gmt80ls), max(raw_vrc01_data$gmt80ls))
vrc01_ic50_range = c(min(raw_vrc01_data$gmt50ls), max(raw_vrc01_data$gmt50ls))

vrc01_pkpd_data = read_csv(here('output/pk-modeling/sim-pk-withVLneut.csv'), col_types = cols())
vrc01_first_pos = subset(vrc01_pkpd_data, first_positive) %>%
  mutate(PT80 = centr_conc/gmt80ls)

theta_vl_vrc01 = theta_vl_placebo

```

```{r pd-setup}

vl_summary = read_csv(clean_data_here("final_vl_summary.csv"), col_types = cols())

#1.28, use 1.25
mean_hill = mean(vl_summary$hill_ls)

amp_theta_pd = tibble(IC80 = NA, rho = 10^final_pop_ests_wide$lrho_pop, h = 1.25)

kable(amp_theta_pd, caption = "AMP PD parameters") %>% kable_styling(full_width = F)

```

```{r pk-setup}
pk_parms = tribble(
  ~parameter, ~value,
  "Cl", 0.575,
  "V1", 3.532,
  "V2", 4.749,
  "Q", 0.676
)

kable(pk_parms, caption = "AMP PK parameters") %>% kable_styling(full_width = F)

theta_pk = create_theta(pk_parms)

conc_range = range(vrc01_pkpd_data$centr_conc, na.rm = T)

```

# Placebo 

the effect on log10 p:

$$lp = lp_0 + \beta_{IC80} (log_{10}IC80 - 0.638)$$

where 0.638 is the mean of the log10 IC80.

```{r}

# overlay the individual draws, at peak I

tibble(
  ic80 = seq(min(raw_placebo_data$gmt80ls), max(raw_placebo_data$gmt80ls), by = 0.1),
  p = burst_effect(ic80)
) %>%
  ggplot(aes(x = ic80, y = p)) +
  geom_line() +
  scale_x_log10() +
  labs(y = "Avg. infectious virions / cell day")

```


```{r placebo-indirect}
ic80_range_burst_size = c(round(placebo_ic80_range, 2), 1, 10)

burst_size_models = map_df(ic80_range_burst_size, function(i){
  theta_vl = theta_vl_placebo
  theta_vl$p = burst_effect(i, pop_parms = placebo_pop_wide)
  
  run_pkpd_models(
        0:100,
        holte_pkpd_model,
        theta_pk = theta_pk,
        theta_vl = theta_vl,
        theta_pd = tibble(IC80 = 4, rho = Inf, h = 1),
        infection_time = 0,
        initial_pk_dose = 0
      ) %>% 
        mutate(isolate_ic80 = i, burst_size = theta_vl$p)
  })

placebo_adj_vl_pl = ggplot(burst_size_models, aes(x = time/7, y = log10V, 
                                                  colour = factor(isolate_ic80))) +
  geom_line() +
  labs(x = "Weeks post-acquisition", y = "Placebo-population log10 VL",
       colour = "VRC01 IC80 of acquired isolate") +
  scale_color_viridis_d() +
  scale_y_continuous(breaks = -1:8)

placebo_adj_vl_pl
# inset the ic80 vs pi plot # 2 x 2
# 4 x 3

```

# VRC01 analysis

## Dose-response

Seems very minimal in the range of the data

```{r dr-setup}
id80_max = conc_range[2]/placebo_ic80_range[1]
id50_max = conc_range[2]/placebo_ic50_range[1]

id50_max_vrc01 = conc_range[2]/vrc01_ic50_range[1]
id80_max_vrc01 = 4^(-1/amp_theta_pd$h) * id50_max_vrc01

dose_response = tibble(
  #id80 = seq(id80_range[1], id80_range[2], by = 0.1),
  id50 = 10^seq(-1, log10(id50_max), by = 0.1),
  PT80 = 4^(-1/amp_theta_pd$h) * id50,
  in_vitro = 1/(1+(id50)^(-amp_theta_pd$h)),
  in_vivo = 1/(1+(id50/amp_theta_pd$rho)^(-amp_theta_pd$h)),
  pe = predict(empirical_amp_fit, data.frame(PT80 = PT80))
)
  
```

```{r dr-plt}
pkpd_dr_pl = dose_response %>%
  gather("type", "neutralization", in_vitro, in_vivo) %>%
  mutate(iip = -log10(1 - neutralization)) %>%
  gather("outcome", "value", neutralization, iip, factor_key = T) %>%
  ggplot(aes(x = PT80, y = value)) +
  geom_line(aes(colour = type)) +
  scale_x_log10() +
  annotate("rect", xmin = id80_max_vrc01, xmax = Inf, ymin = -Inf, ymax = Inf, alpha =0.5, 
           fill = "gray") +
  annotate("text", label = 'placebo\nIC80\nonly', x = 7000, y = 0.65) +
  labs(y = "", colour = "VRC01 PKPD Activity", x = "Theoretical AMP PT80 Range") +
  geom_vline(data = pt80_efficacy, aes(xintercept = PT80), linetype = "dashed") +
  geom_text(data = pt80_efficacy, aes(label = PE_pct), angle = 90, y = 0.65, vjust = 0) +
  facet_wrap(~outcome, scales = "free_y", nrow = 1, strip.position = "left") +
  strip_as_yaxis

pkpd_dr_pl

```


```{r, eval = F}

pkpd_dr_pl_alt = dose_response %>%
  gather("type", "neutralization", in_vitro, in_vivo, pe) %>%
  mutate(
    iip = -log10(1 - neutralization),
    percentage = neutralization * 100
    ) %>%
  gather("outcome", "value", percentage, iip, factor_key = T) %>%
  mutate(
    activity = factor(type, levels = c("in_vitro", "in_vivo", "pe"),
                      labels = c("Neutralization (in vitro)", "PKPD (in vivo)", "PE"))
  ) %>%
  ggplot(aes(x = PT80, y = value)) +
  geom_line(aes(colour = activity)) +
  scale_x_log10() +
  annotate("rect", xmin = id80_max_vrc01, xmax = Inf, ymin = -Inf, ymax = Inf, alpha =0.5, 
           fill = "gray") +
  annotate("text", label = 'placebo\nIC80\nonly', x = 7000, y = 5) +
  labs(y = "", colour = "VRC01 Activity", x = "Theoretical AMP PT80 Range") +
  facet_wrap(~outcome, scales = "free_y", nrow = 1, strip.position = "left") +
  strip_as_yaxis

pkpd_dr_pl_alt

```

## PKPD Effect on Viral Load

```{r setup}

id80_range = c(0, 10, 200, 1000, 2500, 10000)

```

In the context of AMP, need to show the effect of titer on viral load,

- show how R0 and titer are related at time of infection
- within AMP pick different levels of titer, using VRC01 PK with and without burst size adjustment

### R0 vs. titer

This does not adjust for burst size. If this wer eincluded, at a fixed titer, lower IC80 would have lower extinction probability (higher R0).

```{r r0-titer}

# 0.329877
# theta_vl_vrc01$Bt0 * 1/(1 + (1000/(0.329877 * amp_theta_pd$rho))^amp_theta_pd$h)
# 4.995823e-06 * theta_vl_vrc01$p * theta_vl_vrc01$aS / (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23)


r0_titer_sim = tibble(
  PT80 = 1:10000,
  ID50 = PT80 * 4 ^(1/amp_theta_pd$h),
  amp_rho = amp_theta_pd$rho,
  pegu_rho = 100
  ) %>%
  gather("potency_reduction", "rho", amp_rho, pegu_rho) %>%
  mutate(
    beta_neut = beta_titer_neut(ID50/rho, amp_theta_pd$h),
    R0 = beta_neut*theta_vl_vrc01$Bt0 * theta_vl_vrc01$p * theta_vl_vrc01$aS /
        (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23),
    extinction_prob = pmin(1, 1/R0)
    ) 

r0_titer_sim %>%
  ggplot(aes(x = PT80, y = R0)) +
  geom_line(aes(colour = potency_reduction)) +  
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_x_log10()


r0_titer_sim %>%
  ggplot(aes(x = PT80, y = extinction_prob)) +
  geom_line(aes(colour = potency_reduction)) +  
  geom_vline(data = pt80_efficacy, aes(xintercept = PT80), linetype = "dashed") +
  geom_text(data = pt80_efficacy, aes(label = PE_pct), angle = 90, y = 0.65, vjust = 0) +
  scale_x_log10()


```

### PKPD Trajectories with fixed PT80

Ie, if there were constant levels of bnab with fixed PT80. (no indirect effect)

```{r amp-pkpd-pt80-trajectories}

amp_titer_models = map_df(id80_range, function(pt80){
  amp_theta_pd$IC80 = 1
  initial_dose = pt80 * theta_pk$V1
  theta_pk$Cl = 0
  theta_pk$Q = 0
  
  run_pkpd_models(
    c(seq(0, 1, by = 0.1), 2:100),
    holte_pkpd_model,
    theta_pk = theta_pk,
    theta_vl = theta_vl_vrc01,
    theta_pd = amp_theta_pd,
    infection_time = 0,
    initial_pk_dose = initial_dose
  ) %>%
    mutate(
      IC80 = 1,
      PT80 = pt80,
      min_VL = min(log10V)
    )
})

ggplot(amp_titer_models, aes(x = time/7, y = pmax(0.01, centr/IC80), colour = factor(PT80))) +
  scale_y_log10("PT80", breaks = c(0.01, 10, 100, 1000), 
                labels = c(0, 10, 100, 1000)) +
  geom_line() +
  labs( x = "Weeks relative to crossing detection threshold", colour = "PT80 at infection")

amp_titer_models %>%
  ggplot(aes(x = time/7, y = pmax(log10(20), log10V), 
                             colour = factor(PT80))) +
  scale_linetype_discrete(breaks = T, labels = "Potential suppression") +
  scale_y_continuous(breaks = c(log10(20), 2:7), labels = c("< LoD", 2:7)) +
  labs(y = "Log10 VL", x = "Weeks", colour = "Constant PT80", 
       linetype = "") +
  geom_line() 

```

### PKPD Trajectories with PK

Direct effect only. PK follows VRC01. Does not currently show AMP dose but a PT80 of 1000 would be like 
[1000 ug/mL] / IC80 of 1, roughly the peak of high dose

```{r amp-pkpd-trajectories}

amp_pkpd_models = map_df(id80_range, function(pt80){
  amp_theta_pd$IC80 = 1
  initial_dose = pt80 * theta_pk$V1
  
  run_pkpd_models(
    c(seq(0, 1, by = 0.1), 2:100),
    holte_pkpd_model,
    theta_pk = theta_pk,
    theta_vl = theta_vl_vrc01,
    theta_pd = amp_theta_pd,
    infection_time = 0,
    initial_pk_dose = initial_dose
  ) %>%
    mutate(
      IC80 = 1,
      PT80 = pt80,
      min_VL = min(log10V),
      min_detection_time = min(time[log10V > log10(20)]),
      R0 = theta_vl_vrc01$Bt0 * theta_vl_vrc01$p * theta_vl_vrc01$aS /
        (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23),
      R0_adj = Bt0_adj[1] * theta_vl_vrc01$p * theta_vl_vrc01$aS /
        (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23)
    )
})

ggplot(amp_pkpd_models, aes(x = time, y = centr/IC80, colour = factor(PT80))) +
  scale_y_log10("PT80", breaks = c(0.1, 10, 1000)) +
  geom_line()

# ggplot(amp_pkpd_models, aes(x = (time - min_detection_time)/7, y = centr/IC80, colour = factor(PT80))) +
#   scale_y_log10("PT80", breaks = c(0.1, 10, 1000)) +
#   geom_line() +
#   labs( x = "Weeks relative to crossing detection threshold", colour = "PT80 at infection")

amp_pkpd_models %>%
  mutate(extinction = min_VL < -2) %>%
  ggplot(aes(x = time/7, y = pmax(-2, log10V), 
                             colour = factor(PT80),
                             linetype = factor(extinction))) +
  scale_linetype_discrete(breaks = T, labels = "Potential suppression") +
  labs(y = "Log10 VL", x = "Weeks", colour = "PT80 at infection", 
       linetype = "") +
  geom_line() 
# 
# 
# amp_pkpd_models %>%
#   mutate(extinction = min_VL < -2) %>%
#   ggplot(aes(x = (time - min_detection_time)/7, y = pmax(-2, log10V), 
#                              colour = factor(PT80),
#                              linetype = factor(extinction))) +
#   scale_linetype_discrete(breaks = T, labels = "Potential suppression") +
#   scale_y_continuous(breaks = -1:7) +
#   labs(y = "Log10 VL", x = "Weeks relative to crossing detection threshold", colour = "PT80 at infection", 
#        linetype = "") +
#   geom_line() 

```

### PKPD Trajectories with indirect effects with PK

The PK of the ID80 titers are fixed, but they are calculated using different input IC80 because that affects burst size here.

```{r pkpd-indirect}

ic80_range = c(0.1, 1, 10)

amp_pkpd_indirect_models = map_df(id80_range, function(pt80){
  map_df(ic80_range, function(ic80){
    
    amp_theta_pd$IC80 = ic80
    theta_vl_vrc01$p = burst_effect(ic80, pop_parms = theta_vl_vrc01)
    
    initial_dose = pt80 * ic80 * theta_pk$V1
  
    run_pkpd_models(
      c(seq(0, 1, by = 0.1), 2:100),
      holte_pkpd_model,
      theta_pk = theta_pk,
      theta_vl = theta_vl_vrc01,
      theta_pd = amp_theta_pd,
      infection_time = 0,
      initial_pk_dose = initial_dose
    ) %>%
      mutate(
        IC80 = ic80,
        IC80_lab = paste("IC80 =", ic80),
        PT80 = pt80,
        min_VL = min(log10V),
        min_detection_time = min(time[log10V > log10(20)]),
        R0 = theta_vl_vrc01$Bt0 * theta_vl_vrc01$p * theta_vl_vrc01$aS /
          (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23),
        R0_adj = Bt0_adj[1] * theta_vl_vrc01$p * theta_vl_vrc01$aS /
          (theta_vl_vrc01$dS * theta_vl_vrc01$dI * 23)
      )
  })
})

ggplot(amp_pkpd_indirect_models, aes(x = time, y = centr/IC80, colour = factor(PT80))) +
  scale_y_log10("PT80", breaks = c(0.1, 10, 1000)) +
  facet_wrap(~IC80_lab)+
  geom_line()

amp_pkpd_indirect_models %>%
  mutate(extinction = min_VL < -2) %>%
  ggplot(aes(x = time/7, y = pmax(-2, log10V), 
                             colour = factor(PT80),
                             linetype = factor(extinction))) +
  scale_linetype_discrete(breaks = T, labels = "Potential suppression") +
  labs(y = "Log10 VL", x = "Weeks", colour = "PT80 at infection", 
       linetype = "") +
    facet_wrap(~IC80_lab)+
  geom_line() 

amp_pkpd_indirect_models %>%
  mutate(
    extinction = min_VL < -2,
    PT80_lab = paste("Infection PT80 =", PT80)) %>%
  ggplot(aes(x = time/7, y = pmax(-2, log10V), 
                             colour = factor(IC80),
                             linetype = factor(extinction))) +
  scale_linetype_discrete(breaks = T, labels = "Potential suppression") +
  labs(y = "Log10 VL", x = "Weeks", colour = "IC80 at infection", 
       linetype = "") +
    facet_wrap(~PT80_lab)+
  geom_line() 

```

### Simulated PKPD AMP Summary Measures

Attempt to recreate first positive.

```{r amp-pkpd-sim-summary}

calc_delay = 1
auc_end = 90
fine_id80_range = 10^seq(-0.25, log10(id80_max_vrc01), by = 0.1)

amp_titer_vl_summary = map_df(fine_id80_range, function(pt80){
  amp_theta_pd$IC80 = 1
  initial_dose = pt80 * theta_pk$V1
  
  run_pkpd_models(
    c(seq(0, 20, by = 0.1), 21:150),
    holte_pkpd_model,
    theta_pk = theta_pk,
    theta_vl = theta_vl_vrc01,
    theta_pd = amp_theta_pd,
    infection_time = 0,
    initial_pk_dose = initial_dose
  ) %>%
    mutate(
      IC80 = 1,
      PT80 = pt80,
      min_VL = min(log10V),
      min_detection_time = min(time[log10V > log10(20)]),
      PT80_detection = centr[min_detection_time],
      PT80_2wks = 10^mean(log10(centr[time >= min_detection_time & time < (min_detection_time + 14)]))
    )
}) %>%
  group_by(PT80, PT80_detection, PT80_2wks, min_detection_time) %>%
  summarize(
    avg_2wk = mean(log10V[time >= calc_delay & time <= 14+calc_delay]),
    geo_auc_3mo = tail(pkr::AUC(time[time >= calc_delay & time <= auc_end + calc_delay], 
                                log10V[time >= calc_delay & time <= auc_end + calc_delay])[,1], 1)/auc_end,
    avg_2wk_adj = mean(log10V[time >= min_detection_time & time < (min_detection_time + 14)]),
    geo_auc_3mo_adj = tail(pkr::AUC(time[time >= min_detection_time &
                                       time <= auc_end + min_detection_time], 
                                log10V[time >= min_detection_time &
                                       time <= auc_end + min_detection_time])[,1], 1)/auc_end,
    .groups = "drop"
  )

amp_titer_vl_summary %>%
  gather("measure", "value", avg_2wk_adj) %>%
  ggplot(aes(x = PT80_2wks, y = value)) +
  geom_line() +
  geom_point(data = vrc01_first_pos, aes(x = pmax(0.1, PT80), y = log10vl),
             alpha = 0.5) +
scale_x_log10() +
  labs(y = "predicted first positive", x = "predicted PT80 at first positive") 



```

